---
title: "Larval Analysis"
author: "Meghan M. Shea"
date: "`r Sys.Date()`"  
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.height = 7, 
  fig.width = 6.5, 
  echo = TRUE,
  fig.align = 'center',
  cache = FALSE
)

options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

# Set Up

### Load Libraries

```{r Packages, include=TRUE}
# Packages used
packages = c("tidyverse", "knitr", "kableExtra", "ggpmisc")

## Load or install&load
package.check <- lapply(packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

```

### Import Data
```{r Data Import, include=TRUE}

ddPCR_data <- read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_1Feb2026_LarvalSamples.csv", stringsAsFactors = FALSE)
ddPCR_data_diluted <- read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_6Feb2026_LarvalSamplesDiluted.csv", stringsAsFactors = FALSE)

```


# Data Cleaning & Quality Control

First, we want to confirm our reaction details

```{r Reaction details, include=TRUE}

rxn_volume = 20
volume_post_DNA_extraction = 100
template_volume = 5
volume_filtered = 1 

```


Then, we want to check that our NTC and positive control are as expected and that all samples have generated sufficient (>10000 droplets)
```{r Control Check, include=TRUE}

# Create a list of datasets to check
datasets_to_check <- list(
  "ddPCR_data" = ddPCR_data,
  "ddPCR_data_diluted" = ddPCR_data_diluted
)

# Loop through each dataset
for (dataset_name in names(datasets_to_check)) {
  
  cat("\n========================================\n")
  cat("Checking:", dataset_name, "\n")
  cat("========================================\n\n")
  
  current_data <- datasets_to_check[[dataset_name]]
  
  # Check if all NTC samples have concentration of 0
  ntc_check <- current_data %>%
    filter(Sample.Type == "NTC") %>%
    filter(Conc.copies.µL. != 0)
  
  # View any NTC samples that don't have 0 concentration
  if (nrow(ntc_check) > 0) {
    print("WARNING: The following NTC samples have non-zero concentrations:")
    print(ntc_check)
  } else {
    print("✓ All NTC samples have concentration = 0")
  }
  
  
  # Check if positive control sample(s) are close (within 20%) to the expected concentration
  
  # Filter for gBlock samples and extract the number
  gblock_check <- current_data %>%
    filter(grepl("gBlock", Sample.description.1)) %>%
    mutate(
      expected_conc = as.numeric(str_extract(Sample.description.1, "\\d+")),
      adjusted_conc = Conc.copies.µL. * rxn_volume/template_volume,
      percent_diff = abs((adjusted_conc - expected_conc) / expected_conc) * 100,
      within_20pct = percent_diff <= 20
    )
  
  # Display samples that fail the 20% threshold
  failing_samples <- gblock_check %>%
    filter(!within_20pct)
  
  if (nrow(failing_samples) > 0) {
    print("WARNING: The following gBlock samples are NOT within 20% of expected concentration:")
    print(failing_samples %>% select(Sample.description.1, expected_conc, adjusted_conc, percent_diff))
  } else {
    print("✓ All gBlock samples are within 20% of expected concentration")
  }
  
  
  # Check if all samples have sufficient (>10000) droplets
  low_droplets <- current_data %>%
    filter(Accepted.Droplets < 10000) 
  
  # Display results
  if (nrow(low_droplets) > 0) {
    print("WARNING: The following samples have fewer than 10,000 accepted droplets:")
    print(low_droplets)
  } else {
    print("✓ All samples have at least 10,000 accepted droplets")
  }
  
  cat("\n")
}

```




Then, we want to convert our ddPCR output in copies per ul into copies per L of water filtered (or just copies/individual, for larval body samples), for both our diluted and undilted samples
```{r Concentration Conversion, include=TRUE}

dilution_factor = 20

#Samples where individual larva were added to filters

ddPCR_larval_bodies <- ddPCR_data_diluted %>% filter(Sample.Type == "L")
ddPCR_larval_bodies$copies.per.L = (ddPCR_larval_bodies$Conc.copies.µL. * rxn_volume * dilution_factor * (volume_post_DNA_extraction/template_volume)/ddPCR_larval_bodies$Larvae.Number)


#Samples where water samples were collected from larval mesocosms

ddPCR_larval_water <- ddPCR_data_diluted %>% filter(Sample.Type == "LW")
  
ddPCR_larval_water$copies.per.L = (ddPCR_larval_water$Conc.copies.µL. * rxn_volume * dilution_factor * (volume_post_DNA_extraction/template_volume))/(volume_filtered)

ddPCR_larval_water$PoissonConfMax.per.L = (ddPCR_larval_water$PoissonConfMax * rxn_volume * dilution_factor * (volume_post_DNA_extraction/template_volume))/(volume_filtered)
ddPCR_larval_water$PoissonConfMin.per.L = (ddPCR_larval_water$PoissonConfMin * rxn_volume * dilution_factor * (volume_post_DNA_extraction/template_volume))/(volume_filtered)


ddPCR_larval_water_undiluted <- ddPCR_data %>% filter(Sample.Type == "LW")
ddPCR_larval_water_undiluted$copies.per.L = (ddPCR_larval_water_undiluted$Conc.copies.µL. * rxn_volume * (volume_post_DNA_extraction/template_volume))/(volume_filtered)
ddPCR_larval_water_undiluted$PoissonConfMax.per.L = (ddPCR_larval_water_undiluted$PoissonConfMax * rxn_volume * (volume_post_DNA_extraction/template_volume))/(volume_filtered)
ddPCR_larval_water_undiluted$PoissonConfMin.per.L = (ddPCR_larval_water_undiluted$PoissonConfMin * rxn_volume * (volume_post_DNA_extraction/template_volume))/(volume_filtered)


existing_samples <- ddPCR_larval_water %>% 
  pull(Sample.description.1) %>% 
  unique()

# Combine the diluted data with the filtered undiluted data
ddPCR_larval_water <- bind_rows(ddPCR_larval_water, 
                        ddPCR_larval_water_undiluted %>%
  filter(!Sample.description.1 %in% existing_samples))


```

# Analysis

Now, we want to summarize our data across sample types. First, we'll look at our samples where we pipetted individual larvae onto our Sterivex filters. Here, we're interested in the number of copies generated per individual larvae.
```{r Larval Bodies, include=TRUE}
summary_table_1 <- ddPCR_larval_bodies %>%
  summarise(
    n = n(),
    mean_copies_per_L = mean(copies.per.L, na.rm = TRUE),
    sd_copies_per_L = sd(copies.per.L, na.rm = TRUE),
    min_copies_per_L = min(copies.per.L, na.rm = TRUE),
    max_copies_per_L = max(copies.per.L, na.rm = TRUE),
    .groups = "drop"
  )

summary_table_1 %>%
  kable(col.names = c("N", "Mean", "SD", "Min", "Max"),
        caption = "eDNA generated directly by larval bodies (copies/individual)",
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE,
                position = "left") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#2C3E50")


```


Now, we'll look at our samples generated from mesocosms containing between 1 and 10,000 larva.

```{r Larval Water, include=TRUE}
summary_table_2 <- ddPCR_larval_water %>%
  group_by(Larvae.Number) %>%
  summarise(
    n = n(),
    mean_copies_per_L = mean(copies.per.L, na.rm = TRUE),
    sd_copies_per_L = sd(copies.per.L, na.rm = TRUE),
    min_copies_per_L = min(copies.per.L, na.rm = TRUE),
    max_copies_per_L = max(copies.per.L, na.rm = TRUE),
    .groups = "drop"
  )

summary_table_2 %>%
  kable(col.names = c("Number of Larva", "N", "Mean", "SD", "Min", "Max"),
        caption = "eDNA present in mesocosms (copies/L)",
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE,
                position = "left") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#2C3E50")




ddPCR_larval_water <- ddPCR_larval_water %>%
  mutate(copies.per.L.log = log10(copies.per.L),
         Larvae.Number.log = log10(Larvae.Number),
         PoissonConfMax.per.L.log = log10(PoissonConfMax.per.L),
          PoissonConfMin.per.L.log = log10(PoissonConfMin.per.L) )


ggplot(ddPCR_larval_water, aes(x = Larvae.Number.log, y = copies.per.L.log)) +
  geom_errorbar(aes(ymin = PoissonConfMin.per.L.log, 
                    ymax = PoissonConfMax.per.L.log), 
                width = .1, color = "Black") +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), 
                                 after_stat(rr.label),
                                 after_stat(p.value.label),
                                 sep = "~~~"))) +
  labs(x = "Number of Larvae in 10L Container - log10(individuals)",
       y = "Pycnopodia eDNA - log10(Copies/L)") +
  theme_classic(base_size = 12)

ggsave("Fig_6_larval_concentrations.png", 
       width = 7, height = 4, units = "in", 
       dpi = 300)

```





