---
title: "Healthy vs. Wasting Analysis"
author: "Meghan M. Shea"
date: "`r Sys.Date()`"  
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.height = 7, 
  fig.width = 6.5, 
  echo = TRUE,
  fig.align = 'center',
  cache = FALSE
)

options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

# Set Up

### Load Libraries

```{r Packages, include=TRUE}
# Packages used
packages = c("tidyverse", "knitr", "kableExtra")

## Load or install&load
package.check <- lapply(packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

```

### Import Data
```{r Data Import, include=TRUE}

ddPCR_data <- read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_22Jan2026_SonomaFieldSamples.csv", stringsAsFactors = FALSE)

```


# Data Cleaning & Quality Control

First, we want to confirm our reaction details

```{r Reaction details, include=TRUE}

rxn_volume = 20
volume_post_DNA_extraction = 100
template_volume = 5

```


Then, we want to check that our NTC and positive control are as expected and that all samples have generated sufficient (>10000 droplets)
```{r Control Check, include=TRUE}

# Check if all NTC samples have concentration of 0
ntc_check <- ddPCR_data %>%
  filter(Sample.Type == "NTC") %>%
  filter(Conc.copies.µL. != 0)

# View any NTC samples that don't have 0 concentration
if (nrow(ntc_check) > 0) {
  print("WARNING: The following NTC samples have non-zero concentrations:")
  print(ntc_check)
} else {
  print("✓ All NTC samples have concentration = 0")
}


#Check if positive control sample(s) are close (within 20%) to the expected concentration

# Filter for gBlock samples and extract the number
gblock_check <- ddPCR_data %>%
  filter(grepl("gBlock", Sample.description.1)) %>%
  mutate(
    expected_conc = as.numeric(str_extract(Sample.description.1, "\\d+")),
    adjusted_conc = Conc.copies.µL. * rxn_volume/template_volume,
    percent_diff = abs((adjusted_conc - expected_conc) / expected_conc) * 100,
    within_20pct = percent_diff <= 20
  )

# Display samples that fail the 10% threshold
failing_samples <- gblock_check %>%
  filter(!within_20pct)

if (nrow(failing_samples) > 0) {
  print("WARNING: The following gBlock samples are NOT within 10% of expected concentration:")
  print(failing_samples %>% select(Sample.description.1, expected_conc, adjusted_conc, percent_diff))
} else {
  print("✓ All gBlock samples are within 10% of expected concentration")
}


# Check if all samples have sufficient (>10000) droplets
low_droplets <- ddPCR_data %>%
  filter(Accepted.Droplets < 10000) 

# Display results
if (nrow(low_droplets) > 0) {
  print("WARNING: The following samples have fewer than 10,000 accepted droplets:")
  print(low_droplets)
} else {
  print("✓ All samples have at least 10,000 accepted droplets")
}

```

Then, we want to convert our ddPCR output in copies per ul into copies per L of water filtered. 
```{r Concentration Conversion, include=TRUE}

ddPCR_data$copies.per.L = (ddPCR_data$Conc.copies.µL. * rxn_volume * (volume_post_DNA_extraction/template_volume))/(ddPCR_data$Volume.Filtered/1000)

```

# Analysis

Now, we want to summarize our data across sample types. 

```{r Summaries, include=TRUE}
# Filter out PC and NTC samples, then summarize for each combination of sample type & distance
summary_table_1 <- ddPCR_data %>%
  filter(!Sample.Type %in% c("PC", "NTC")) %>%
  group_by(Sample.Type, Distance) %>%
  summarise(
    n = n(),
    mean_copies_per_L = mean(copies.per.L, na.rm = TRUE),
    sd_copies_per_L = sd(copies.per.L, na.rm = TRUE),
    min_copies_per_L = min(copies.per.L, na.rm = TRUE),
    max_copies_per_L = max(copies.per.L, na.rm = TRUE),
    .groups = "drop"
  )

summary_table_1 %>%
  kable(col.names = c("Sample Type", "Distance", "N", "Mean", "SD", "Min", "Max"),
        caption = "eDNA (Copies per L) by Sample Type and Distance",
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE,
                position = "left") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#2C3E50")

# Summarize across all healthy samples
summary_table_2 <- ddPCR_data %>%
  filter(Sample.Type %in% c("Healthy")) %>%
  summarise(
    n = n(),
    mean_copies_per_L = mean(copies.per.L, na.rm = TRUE),
    sd_copies_per_L = sd(copies.per.L, na.rm = TRUE),
    min_copies_per_L = min(copies.per.L, na.rm = TRUE),
    max_copies_per_L = max(copies.per.L, na.rm = TRUE),
    .groups = "drop"
  )

summary_table_2 %>%
  kable(col.names = c("N", "Mean Copies/L", "SD", "Min", "Max"),
        caption = "eDNA (Copies per L) for All Healthy Samples",
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "left") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#27AE60")

```