---
title: "Pycno Run 5"
author: "Zack Gold"
date: "2025-12-30"
output: pdf_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE,fig.height=7, fig.width=6.5, echo = TRUE)
knitr::opts_chunk$set(fig.pos = '!h')
sink.indicator <- TRUE

```

# Set Up

### Load Libraries

```{r Packages, include=FALSE}
# Packages used
packages = c("tidyverse", 
             "dplyr",
             "ggplot2",
             "knitr",
             "kableExtra",
             "RColorBrewer",
             "readxl",
             "here",
             "Hmisc",
             "scales",
              "ggpmisc",
             "ggpubr",
             "wesanderson",
             "RColorBrewer",
             "randomcoloR",
             "ggdist",
             "gmodels",
             "rcompanion",
             "gghalves",
             "ggbeeswarm",
             "rstatix",
             "multcompView",
             "ggstatsplot",
             "knitr")

## Load or install&load
package.check <- lapply(packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

## Source the functions
source("eLowQuantUser/eLowQuant-Functions-V20210407.R")

```


# Values Set
```{r}
IPC_name = "Moa"
Target_name = "Pycno"

```

# Load Data Pycno Plate Run 5 

```{r}
sample_set_up <- read_excel(here("qPCR_raw_data","2025-12-23_161523_Pycno_5_v3.xls"), sheet="Sample Setup", skip = 46)

amplification_data<- read_excel(here("qPCR_raw_data","2025-12-23_161523_Pycno_5_v3.xls"), sheet="Amplification Data",  skip = 46)

results_data<- read_excel(here("qPCR_raw_data","2025-12-23_161523_Pycno_5_v3.xls"), sheet="Results",  skip = 46)

```

### Merge Tables
```{r}

amplification_data %>% 
  left_join(sample_set_up) -> combined_qPCR_data

```

Now we are ready to start analyzing the data.

First thing we want to do is check to see how our internal positive control is working.

## Moa Map
```{r}

columns = seq(from =1, to=24, by=1)
columns_t= as.tibble(columns) %>%  dplyr::rename("Column"=value)
rows = capitalize(letters[1:16])
rows_t= as.tibble(rows) %>%  dplyr::rename("Row"=value)

cross_join(columns_t, rows_t) -> full_plate

combined_qPCR_data %>% 
  filter(., `Target Name`==IPC_name) %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> moa_samples
  
full_plate %>% 
  left_join(moa_samples) -> moa_plate


palette <- distinctColorPalette(length(moa_plate$`Sample Name` %>%  unique()))

moa_plate %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns , na.value = "black") + ylab("Row") +theme_pubr() +scale_fill_manual(values =palette, na.value = "black") +ggtitle("Moa-IPC Standards & Samples")


```



## Pycno Map
```{r}

combined_qPCR_data %>% 
  filter(., `Target Name`==Target_name) %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> pycno_samples
  
full_plate %>% 
  left_join(pycno_samples) -> pycno_plate

palette <- distinctColorPalette(length(pycno_plate$`Sample Name` %>%  unique()))

pycno_plate %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns ) + ylab("Row") +theme_pubr() +scale_fill_manual(values =palette, na.value = "black") +ggtitle("Pycno Standards & Samples")

```


## Moa Rn
```{r}

combined_qPCR_data %>% 
filter(., str_detect(`Target Name`,IPC_name)) -> moa_standard

moa_standard$Rn %>%  max() -> max_Rn_moa
h_moa= max_Rn_moa*0.25

moa_standard %>% 
 # filter(., `Well Position`=="A24") %>% A24 did not perform well
ggplot(., aes(y=Rn, x=Cycle, color=`Well Position`)) +
  geom_point(alpha=0.2) + geom_hline(yintercept =h_moa) +
  geom_text(aes(0,h_moa,label = round(h_moa,2), vjust = -1),color="black") +theme_bw() +ggtitle("Moa Rn") +theme(legend.position = "blank")


```

There is quite a bit of spread in our IPC. Will remove any samples outside normal distribution.
#### Remove Samples with Anomolous Outlier IPC values
```{r}

moa_standard %>% 
  filter(., Cycle ==40) -> moa_standard_c40

moa_standard_c40 %>% 
 dplyr::summarise(mean_Rn = ci(Rn, confidence = 0.99)[1], 
                      lower_ci_mpg = ci(Rn, confidence = 0.99)[2],
                      upper_ci_mpg = ci(Rn, confidence = 0.99)[3], 
                      sd_Rn = ci (Rn, confidence = 0.99)[4]) -> CI0.95_moa

Sum = groupwiseMean(Rn ~ Cycle, data=moa_standard_c40, conf = 0.99)

moa_standard_c40 %>% 
summarise(
    Q1 = quantile(Rn, 0.25, na.rm = TRUE),
    Q3 = quantile(Rn, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR
  ) -> moa_iqr

moa_standard_c40 %>% 
  mutate(., type=case_when(Rn > moa_iqr$upper_bound ~"Outlier High",
                           Rn < moa_iqr$lower_bound ~"Outlier Low",
                           TRUE~"To Keep")) -> moa_standard_c40

moa_standard_c40 %>% 
  ggplot(., aes(x= Rn, y=Cycle)) +  geom_dots(layout='weave', side = "top", aes(color=type)) + stat_slabinterval(side = "bottom",aes(fill = after_stat(level)))  +scale_fill_brewer(na.translate = FALSE) +theme_pubclean()

moa_standard_c40 %>% 
  filter(., type!="To Keep") -> moa_outliers_to_drop
```

#### Filter Data
```{r}
combined_qPCR_data %>% 
  filter(., `Well Position` != moa_outliers_to_drop$`Well Position`) -> combined_qPCR_data_cleaned

results_data %>% 
    filter(., `Well Position` != moa_outliers_to_drop$`Well Position`) -> results_data_cleaned

```




## Pycno Rn
```{r}
combined_qPCR_data_cleaned %>% 
filter(., str_detect(`Target Name`,Target_name)) %>% 
  filter(., Task=="STANDARD")-> pycno_standard

pycno_standard$Rn %>%  max() -> max_Rn
h= max_Rn*0.25

pycno_standard %>% 
ggplot(., aes(y=Rn, x=Cycle, color=`Sample Name`)) +geom_point(alpha=0.4) +geom_smooth() + geom_hline(yintercept =h) +
  geom_text(aes(0,h,label = round(h,2), vjust = -1),color="black") +theme_bw() + ggtitle("Pycno Rn") +scale_colour_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))


```



Looks great. We except higher concentration standards to surpass the threshold ( here set to 25% of the maximum Rn flourescence value) at fewer cycles and we see that in the data.

## Pycno Ct versus Quantity
```{r}

pycno_standard %>% 
  mutate(., Threshold = if_else(Rn > h, "Above","Below")) -> pycno_standard

pycno_standard %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="first")==1) %>% 
  filter(., Threshold =="Above") -> pycno_standard_above

pycno_standard %>% 
  ungroup() %>% 
  filter(., !`Well Position` %in%  pycno_standard_above$`Well Position`) %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="last")==1) -> pycno_standard_below

bind_rows(pycno_standard_above, pycno_standard_below) -> pycno_standard_threshold



set_breaks = function(limits) {
     seq(limits[1], limits[2], by = 1)
}

pycno_standard_threshold %>% 
  filter(., Cycle < 40 ) %>% 
  ggplot(., aes(x=`Quantity`, y = Cycle)) +geom_count(aes(color = ..n.., size = ..n..))  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct") -> run5_ct_versus_quant

run5_ct_versus_quant
```

```{r}

# View the structure of the generated plot object
# The computed data is stored within the plot object
grob_data_run5 <- ggplot_build(run5_ct_versus_quant)$data[[3]]

# The result is a list; unlist it to get a character vector
coef_char_run5 <- unlist(grob_data_run5$coefs)


standard_curve_coefs_run5 <- enframe(coef_char_run5, name = "Metric", value = "Value") %>% 
 mutate( Metric = recode(Metric,
                             "(Intercept)" = "Intercept",
                             "x" = "Slope"))

# Extract the R^2 value from the 'r.squared' column
r_squared_value_run5 <- grob_data_run5$r.squared

```

```{r}
results_data_cleaned %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task=="STANDARD") %>% 
   mutate(., CT=as.numeric(CT)) -> pycno_results_standard


 pycno_results_standard %>% 
  ggplot(., aes(x=`Quantity`, y = CT)) +geom_point()  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct")
 
```
```{r READIN, include=FALSE}
## READIN paragraph adapted from Merkes, Christopher <cmerkes@usgs.gov>
##   Used only to read in the data.


## Read in your data file (MODIFY FILE NAME AS NEEDED):
results_data_cleaned %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task!="UNKNOWN") %>% 
   mutate(., CT=as.numeric(CT))  %>% 
  mutate(., Lab="NOAA PMEL OME")  %>% 
  dplyr::select(Target=`Target Name`, Lab, Cq=CT, SQ=Quantity) %>% 
  mutate(., TemplateVolume =2,
         ReactionVolume= 10) %>% 
  mutate(., Cq = replace(Cq, Cq>40, NA))-> qPCR_Pycno_5_v3_standards

write.csv(qPCR_Pycno_5_v3_standards, file = here("LOD_LOQ/qPCR_Pycno_5_v3_standards.csv"), row.names = F)
DAT <- qPCR_Pycno_5_v3_standards

```

# FHL Analysis Transplanted Individual Detection


## Merge Sample Metadata
```{r}

sample_metadata<- read_excel(here("qPCR_raw_data","ElectronicArchive_CruiseSampleMetadata_20250516_FHL.xlsx"), sheet="Electronic Archive")

results_data_cleaned %>% 
   filter(.,`Target Name`=="Pycno") %>% 
  filter(., !str_detect(`Task`,"STANDARD")) %>% 
  mutate(., Quantity=if_else(is.na(Quantity), 0, Quantity)) %>% 
  mutate(., Concentration = (Quantity*10*(100/2)/ 1) %>% # Quantity is the concentration generated by the qPCR machine in copies per µL, 10 µL per reaction, 100 µL in the total extraction, 2 µL of template added, and 1 L sampled
  mutate(., Observation = if_else(Quantity ==0, "Not Detected", "Detected")) %>%
  mutate(., `Sample Name`= if_else(str_detect(`Sample Name`,"Negative Control"), "NTC",`Sample Name`)) %>% 
  ungroup() %>% 
  left_join(sample_metadata, by=c("Sample Name"="Sample_Name")) %>% 
  mutate(., `Sample Name` = if_else(`Sample Name`=="Ext_blank_UWFH25","Extraction Blank",`Sample Name`)) %>% 
  mutate(., Negative_control = if_else(is.na(Negative_control), TRUE, Negative_control)) %>% 
  group_by(`Sample Name`) %>% 
  mutate(Technical_Replicate = row_number(`Sample Name`)) %>% 
  mutate(., Station = if_else(is.na(Station) & Negative_control==TRUE | Task=="NTC", `Sample Name`, Station)) %>% 
  mutate(., Name_plotting = str_c(Station," Bottle ", Biological_Replicate," PCR ",Technical_Replicate )) %>% 
  mutate(., Name_plotting = if_else(is.na(Name_plotting),str_c(`Sample Name`, " PCR ",Technical_Replicate),Name_plotting)) %>% 
   mutate(., Name_plotting_b = str_c(Station," Bottle ", Biological_Replicate)) %>% 
  mutate(., Name_plotting_b = if_else(is.na(Name_plotting_b),`Sample Name`,Name_plotting_b)) %>% 
  mutate(., Type = if_else(is.na(Station),"Extraction Blank",Station )) %>% 
   mutate(., Type = if_else(str_detect(Type,".NC"),"Field Blank",Type )) %>% 
  mutate(., Sample_Type = if_else(Negative_control==FALSE,"Sample","Blank")) -> pycno_sample_data_4_plotting
```

```{r}
pycno_sample_data_4_plotting %>% 
     mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  ggplot(., aes(x=Name_plotting_b, y=Concentration, color=Observation)) + geom_count() +
    scale_y_continuous( 
         trans = scales::pseudo_log_trans(base = 10), 
         breaks = c(1,10,1000,10000,100000,1000000), 
         labels = scales::trans_format("log10", scales::math_format(10^.x))
     )    +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=0)) +facet_wrap(Sample_Type~Type,nrow=1, scales="free_x") + ylab("Copies per L") + xlab("Sample") 

  
# pycno_sample_data_4_plotting_run1 %>% 
#   filter(., str_detect(Name_plotting_b,"Aquarium")) %>% 
#   group_by(`Target Name`) %>% 
#   dplyr::summarise(mean(Concentration))
```
### Figure 6a2
```{r}
pycno_sample_data_4_plotting

pycno_sample_data_4_plotting$Type2 <- pycno_sample_data_4_plotting$Type
pycno_sample_data_4_plotting$Type2 <- as.factor(pycno_sample_data_4_plotting$Type2)
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "Aquarium Discharge"] <- "Aquarium\nDischarge"
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "FHL Dock"] <- "FHL\nDock"
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "FHL Intake Pipes"] <- "FHL\nIntake\nPipes"

pycno_sample_data_4_plotting %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., Concentration=if_else(Concentration==0,NA,Concentration) ) %>% 
  mutate(., Plate = "5")->  plate_5_data_analysis

plate_5_data_analysis %>% 
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=log10(Concentration), color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black") + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> a2

a2


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_run5.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```

### Figure s6a2
```{r}
pycno_sample_data_4_plotting

pycno_sample_data_4_plotting$Type2 <- pycno_sample_data_4_plotting$Type
pycno_sample_data_4_plotting$Type2 <- as.factor(pycno_sample_data_4_plotting$Type2)
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "Aquarium Discharge"] <- "Aquarium\nDischarge"
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "FHL Dock"] <- "FHL\nDock"
levels(pycno_sample_data_4_plotting$Type2)[levels(pycno_sample_data_4_plotting$Type2) == "FHL Intake Pipes"] <- "FHL\nIntake\nPipes"

pycno_sample_data_4_plotting %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=CT, color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black",limits = c(20, 40),direction = -1) + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> sa2

sa2


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_Ct_run5.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```

Here we see ~2 million copies per L of Pycnopodia target DNA in the Friday Harbor Aquarium discharge water. The discharge is from a flow through system with 39 individual stars of at least 40 cm in diameter.

We were able to detect the 10 transplanted stars from at least one technical replicate from each of the three surface eDNA samples taken ~5m above the individuals. We only detected a single individual located at 10m depth from one of three surface samples taken at the Friday Harbor Lab intake pipes.

These results demonstrate that the eDNA qPCR assay can detect Pycnopodia from both mesocosms and the field. 


# FHL Analysis Passive Versus Active Filtration
```{r}

results_data_cleaned %>% 
  filter(.,`Target Name`=="Pycno") %>% 
  filter(., !str_detect(`Task`,"STANDARD")) %>% 
    mutate(., `Sample Name`=str_replace(`Sample Name`,",",".")) %>% 
  left_join(sample_metadata, by=c("Sample Name"="Sample_Name")) %>% 
  group_by(`Sample Name`) %>% 
    mutate(., Negative_control = if_else(is.na(Negative_control), TRUE, Negative_control)) %>% 
  mutate(Technical_Replicate = row_number(`Sample Name`)) %>% 
  mutate(., Station = if_else(is.na(Station) & Negative_control=="TRUE" | Task=="NTC", `Sample Name`, Station)) %>% 
  mutate(., Name_plotting = str_c(Station," ",Field_collection_method," ", Biological_Replicate," ",Technical_Replicate )) %>% 
  mutate(., Name_plotting = if_else(is.na(Name_plotting),str_c(`Sample Name`, " ",Technical_Replicate),Name_plotting)) %>% 
   mutate(., Name_plotting_b = str_c(Station," ",Field_collection_method," ", Biological_Replicate)) %>% 
  mutate(., Name_plotting_b = if_else(is.na(Name_plotting_b),`Sample Name`,Name_plotting_b)) %>% 
  mutate(., Quantity=if_else(is.na(Quantity), 0, Quantity)) %>% 
  mutate(., Concentration = (Quantity*10*(100/2)/ 1) %>% # Quantity is the concentration generated by the qPCR machine in copies per µL, 10 µL per reaction, 100 µL in the total extraction, 2 µL of template added, and 1 L sampled
  mutate(., Observation = if_else(Quantity ==0, "Not Detected", "Detected")) %>% 
  mutate(., Type = if_else(is.na(Station),"Extraction Blank",Station )) %>% 
   mutate(., Type = if_else(str_detect(Type,".NC"),"Field Blank",Type )) %>%  mutate(., Sample_Type = if_else(Negative_control==FALSE,"Sample","Blank")) %>% 
  mutate(., Field_collection_method = if_else(Task=="NTC","NTC",Field_collection_method)) %>% 
mutate(., Field_collection_method = if_else(Type=="Extraction Blank","Extraction Blank",Field_collection_method))-> pycno_sample_data_4_plotting
```


```{r}

pycno_sample_data_4_plotting %>% 
  filter(., !str_detect(Name_plotting_b, "FHL")) %>% 
  ggplot(., aes(x=Name_plotting_b, y=Concentration_L, color=Observation)) + geom_count() +
    scale_y_continuous( 
         trans = scales::pseudo_log_trans(base = 10), 
         breaks = c(1,10,1000,10000,100000,1000000,10000000,100000000), 
         labels = scales::trans_format("log10", scales::math_format(10^.x))
     ) +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05)) +facet_wrap(Sample_Type~Field_collection_method, scales="free_x") + ylab("Copies per µL") +xlab("Bottle Replicates") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

```
We had minor contamination in the passive filter. It's 2 orders of magnitude less than any of the samples so not a major concern. Currently not adjusting the values here, but could subtract this value from all observed positive results.


```{r}


pycno_sample_data_4_plotting %>% 
  filter(., !str_detect(Name_plotting_b, "FHL")) %>% 
  filter(., Sample_Type!="Blank") %>% 
  mutate(., Method= case_when(str_detect(Field_collection_method,"Sterivex")~"Active Filtration",str_detect(Field_collection_method,"Smith")~"Active Filtration",
                  TRUE~"Passive Filtration"            )) ->pycno_sample_data_4_plotting_fm
```

### Kruskal Wallace test
```{r}
# Perform one-way ANOVA
kw_test_results <-  kruskal.test(Concentration_L ~ Field_collection_method, pycno_sample_data_4_plotting_fm)

# Perform post-hoc Tukey HSD test
pairwise.wilcox.test(pycno_sample_data_4_plotting_fm$Concentration_L, pycno_sample_data_4_plotting_fm$Field_collection_method,
                 p.adjust.method = "bonferroni") -> pairwise_results

full_p_matrix <- fullPTable(pairwise_results$p.value)
multcompLetters(full_p_matrix) -> cld

cld_tibble_optionA <- as_tibble(cld$Letters, rownames = "group") %>% 
  dplyr::select(Field_collection_method=group, Letters=value)

pairwise_results
```


```{r}  

pycno_sample_data_4_plotting_fm$Field_collection_method2 <- pycno_sample_data_4_plotting_fm$Field_collection_method
pycno_sample_data_4_plotting_fm$Field_collection_method2 <- as.factor(pycno_sample_data_4_plotting_fm$Field_collection_method2)
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Kringle Gauze"] <- "Kringle\nGauze"
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Vivecare gauze"] <- "Vivecare\nGauze"
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Tische Environmental Filter"] <- "Tische\nEnvironmental\nFilter"
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Paint Roller"] <- "Paint\nRoller"
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Sterivex 0.22µm"] <- "Sterivex\n0.22µm"
levels(pycno_sample_data_4_plotting_fm$Field_collection_method2)[levels(pycno_sample_data_4_plotting_fm$Field_collection_method2) == "Smith root 0.22 µm"] <- "Smith\nRoot\n0.22 µm"


pycno_sample_data_4_plotting_fm %>% 
  left_join(cld_tibble_optionA) %>% 
  ggplot(., aes(x=Field_collection_method2, y=Concentration_L)) + 
  geom_boxplot(aes(color=Method)) +
  geom_point(aes(color=Method)) +
  scale_y_continuous( 
         trans = scales::pseudo_log_trans(base = 10), 
         breaks = c(1,10,1000,10000,100000,1000000,10000000,100000000), 
         labels = scales::trans_format("log10", scales::math_format(10^.x))
     ) +theme_pubr()  + ylab("Copies per L") +xlab("Sampling Method") +scale_color_manual(values=wes_palette("FrenchDispatch",n=2)) + geom_text(aes(label=Letters, y=10^6)) +theme(axis.title = element_text(size = 16, face = "bold"),  axis.text.x = element_text(size = 10), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) -> a1


a1


ggsave(filename = here("Figures","FHL_mesocosm_passive_active_comparison.png"),
       width=10,
       height = 6,
       dpi = 400,
      units = c("in"))
```


```{r}
pycno_sample_data_4_plotting %>% 
  filter(., !str_detect(Name_plotting_b, "FHL")) %>% 
  filter(., Sample_Type!="Blank") %>% 
  mutate(., Method= case_when(str_detect(Field_collection_method,"Sterivex")~"Active Filtration",str_detect(Field_collection_method,"Smith")~"Active Filtration",
                  TRUE~"Passive Filtration"            )) %>% 
  group_by(Field_collection_method, Method) %>% 
  dplyr::summarize(Max=max(Concentration_L), Min=min(Concentration_L), Mean=mean(Concentration_L), median = median(Concentration_L),
    IQR = IQR(Concentration_L)) %>% left_join(cld_tibble_optionA) %>% kable(digits=2)

```



Here we see that Sterivex 0.22 µm filters captured nearly an order of magnitude more Pycnopodia DNA than either Smith Root active filters or any of the passive filters deployed here. This is notable given that Sterivex have a smaller surface area (10cm^2) than the Smith Root 47mm filter (17.3 cm^2) and suggests that differences in the housing or sample integrity are driving these apparent differences. These results highlight that passive filters can be used for field applications to detect Pycnopodia, albeit at a significant disadvantage in that even the best performing materials captured less than a 10th of the DNA captured by active sterivex filtration methods.



# Load Data Pycno qPCR Plate 1

```{r}
sample_set_up_run1 <- read_excel(here("qPCR_raw_data","2025-05-23_132340_Pycno_test_1.xls"), sheet="Sample Setup", skip = 46)

amplification_data_run1 <- read_excel(here("qPCR_raw_data","2025-05-23_132340_Pycno_test_1.xls"), sheet="Amplification Data",  skip = 46)

results_data_run1<- read_excel(here("qPCR_raw_data","2025-05-23_132340_Pycno_test_1.xls"), sheet="Results",  skip = 46)

```

### Merge Tables
```{r}

amplification_data_run1 %>% 
  left_join(sample_set_up_run1) %>% 
  mutate(., `Sample Name` = if_else(is.na(`Sample Name`),str_c(`Target Name`,"_",Task,"_", Quantity), `Sample Name` )) %>% 
  mutate(., `Sample Name` = if_else(is.na(`Sample Name`),"Negative Control", `Sample Name` ))-> combined_qPCR_data_run1

```

Now we are ready to start analyzing the data.

First thing we want to do is check to see how our internal positive control is working.

## Moa Map
```{r}

columns = seq(from =1, to=24, by=1)
columns_t= as.tibble(columns) %>%  dplyr::rename("Column"=value)
rows = capitalize(letters[1:16])
rows_t= as.tibble(rows) %>%  dplyr::rename("Row"=value)

cross_join(columns_t, rows_t) -> full_plate


combined_qPCR_data_run1 %>% 
  filter(., `Target Name`=="Moa-IPC") %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> moa_samples_run1
  
full_plate %>% 
  left_join(moa_samples_run1) -> moa_plate_run1


moa_plate_run1 %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns ) + ylab("Row") +theme_pubr() +scale_fill_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Chevalier1"))) +ggtitle("Moa-IPC Standards & Samples")


```



## Pycno Map
```{r}

combined_qPCR_data_run1 %>% 
  filter(., `Target Name`=="Pycno") %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> pycno_samples_run1
  
full_plate %>% 
  left_join(pycno_samples_run1) -> pycno_plate_run1


pycno_plate_run1 %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns ) + ylab("Row") +theme_pubr() +scale_fill_manual(values =c(wes_palette("Darjeeling1")[1:2],wes_palette("Darjeeling1")[5],wes_palette("Darjeeling2"), wes_palette("Chevalier1"), brewer.pal(n=7, "Oranges"))) +ggtitle("Pycno Standards & Samples")

```


## Moa Rn
```{r}
combined_qPCR_data_run1 %>% 
filter(., str_detect(`Sample Name`,"Moa-IPC_STANDARD")) -> moa_standard_run1

moa_standard_run1$Rn %>%  max() -> max_Rn_moa_run1
h_moa_run1= max_Rn_moa_run1*0.25

moa_standard_run1 %>% 
ggplot(., aes(y=Rn, x=Cycle, color=`Well Position`)) +
  geom_point(alpha=0.2) + geom_hline(yintercept =h_moa_run1) +
  geom_text(aes(0,h_moa_run1,label = round(h_moa_run1,2), vjust = -1),color="black") +theme_bw() +ggtitle("Moa Rn")


```


There is quite a bit of spread in our IPC. Will remove any samples outside normal distribution.
#### Remove Samples with Anomolous Outlier IPC values
```{r}

moa_standard_run1 %>% 
  filter(., Cycle ==40) -> moa_standard_c40_run1

moa_standard_c40_run1 %>% 
 dplyr::summarise(mean_Rn = ci(Rn, confidence = 0.99)[1], 
                      lower_ci_mpg = ci(Rn, confidence = 0.99)[2],
                      upper_ci_mpg = ci(Rn, confidence = 0.99)[3], 
                      sd_Rn = ci (Rn, confidence = 0.99)[4]) -> CI0.95_moa_run1

Sum_run1 = groupwiseMean(Rn ~ Cycle, data=moa_standard_c40_run1, conf = 0.99)

moa_standard_c40_run1 %>% 
summarise(
    Q1 = quantile(Rn, 0.25, na.rm = TRUE),
    Q3 = quantile(Rn, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR
  ) -> moa_iqr_run1

moa_standard_c40_run1 %>% 
  mutate(., type=case_when(Rn > moa_iqr_run1$upper_bound ~"Outlier High",
                           Rn < moa_iqr_run1$lower_bound ~"Outlier Low",
                           TRUE~"To Keep")) -> moa_standard_c40_run1

moa_standard_c40_run1 %>% 
  ggplot(., aes(x= Rn, y=Cycle)) +  geom_dots(layout='weave', side = "top", aes(color=type)) + stat_slabinterval(side = "bottom",aes(fill = after_stat(level)))  +scale_fill_brewer(na.translate = FALSE) +theme_pubclean()

moa_standard_c40_run1 %>% 
  filter(., type!="To Keep") -> moa_outliers_to_drop_run1
```

#### Filter Data
```{r}
combined_qPCR_data_run1 %>% 
  filter(., `Well Position` != moa_outliers_to_drop_run1$`Well Position`) -> combined_qPCR_data_cleaned_run1


results_data_run1 %>% 
    filter(., `Well Position` != moa_outliers_to_drop$`Well Position`) -> results_data_cleaned_run1
```




## Pycno Rn
```{r}
combined_qPCR_data_run1 %>% 
filter(., str_detect(`Sample Name`,"Pycno_STANDARD")) -> pycno_standard_run1

pycno_standard_run1$Rn %>%  max() -> max_Rn_run1
h_run1= 0.8

pycno_standard_run1 %>% 
ggplot(., aes(y=Rn, x=Cycle, color=`Sample Name`)) +geom_point(alpha=0.4) + geom_hline(yintercept =h_run1) +
  geom_text(aes(0,h_run1,label = round(h_run1,2), vjust = -1),color="black") +theme_bw() + ggtitle("Pycno Rn") +scale_colour_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))


```

#### Pycno Ct versus Quantity
```{r}

pycno_standard_run1 %>% 
  mutate(., Threshold = if_else(Rn > h_run1, "Above","Below")) -> pycno_standard_run1

pycno_standard_run1 %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="first")==1) %>% 
  filter(., Threshold =="Above") -> pycno_standard_above_run1

pycno_standard_run1 %>% 
  ungroup() %>% 
  filter(., !`Well Position` %in%  pycno_standard_above_run1$`Well Position`) %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="last")==1) -> pycno_standard_below_run1

bind_rows(pycno_standard_above_run1, pycno_standard_below_run1) -> pycno_standard_threshold_run1



set_breaks = function(limits) {
     seq(limits[1], limits[2], by = 1)
}

pycno_standard_threshold_run1 %>% 
    filter(., Cycle < 40 ) %>% 
  ggplot(., aes(x=`Quantity`, y = Cycle)) +geom_count(aes(color = ..n.., size = ..n..))  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct") -> run1_ct_versus_quant


run1_ct_versus_quant
```

```{r}

# View the structure of the generated plot object
# The computed data is stored within the plot object
grob_data_run1 <- ggplot_build(run1_ct_versus_quant)$data[[3]]

# The result is a list; unlist it to get a character vector
coef_char_run1 <- unlist(grob_data_run1$coefs)


standard_curve_coefs_run1 <- enframe(coef_char_run1, name = "Metric", value = "Value") %>% 
 mutate( Metric = recode(Metric,
                             "(Intercept)" = "Intercept",
                             "x" = "Slope"))

# Extract the R^2 value from the 'r.squared' column
r_squared_value_run1 <- grob_data_run1$r.squared

```

```{r}
results_data_cleaned_run1 %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task=="STANDARD") %>% 
   mutate(., CT=as.numeric(CT)) -> pycno_results_standard_run1


 pycno_results_standard_run1 %>% 
  ggplot(., aes(x=`Quantity`, y = CT)) +geom_point()  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct")
 
```



```{r}
results_data_cleaned_run1 %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task!="UNKNOWN") %>% 
   mutate(., CT=as.numeric(CT))  %>% 
  mutate(., Lab="NOAA PMEL OME") %>% 
  dplyr::select(Target=`Target Name`, Lab, Cq=CT, SQ=Quantity) %>% 
   mutate(., TemplateVolume =2,
         ReactionVolume= 10) %>% 
  mutate(., Cq = replace(Cq, Cq>40, NA))-> qPCR_Pycno_1_standards

write.csv(qPCR_Pycno_1_standards, file = here("LOD_LOQ/qPCR_Pycno_1_standards.csv"), row.names = F)
```

## FHL Analysis Transplanted Individual Detection

## Merge Sample Metadata
```{r}

sample_metadata<- read_excel(here("qPCR_raw_data","ElectronicArchive_CruiseSampleMetadata_20250516_FHL.xlsx"), sheet="Electronic Archive")

results_data_cleaned_run1 %>% 
   filter(.,`Target Name`=="Pycno") %>% 
  filter(., !str_detect(`Task`,"STANDARD")) %>% 
    mutate(., Quantity=if_else(is.na(Quantity), 0, Quantity)) %>% 
  mutate(., Concentration = (Quantity*10*(100/2)/ 1) %>% # Quantity is the concentration generated by the qPCR machine in copies per µL, 10 µL per reaction, 100 µL in the total extraction, 2 µL of template added, and 1 L sampled
  mutate(., Observation = if_else(Quantity ==0, "Not Detected", "Detected")) %>%
  mutate(., `Sample Name`= if_else(str_detect(`Sample Name`,"Negative Control"), "NTC",`Sample Name`)) %>% 
  ungroup() %>% 
  left_join(sample_metadata, by=c("Sample Name"="Field_Name")) %>% 
  mutate(., `Sample Name` = if_else(`Sample Name`=="Ext_blank_UWFH25","Extraction Blank",`Sample Name`)) %>% 
  mutate(., Negative_control = if_else(is.na(Negative_control), TRUE, Negative_control)) %>% 
  group_by(`Sample Name`) %>% 
  mutate(Technical_Replicate = row_number(`Sample Name`)) %>% 
  mutate(., Station = if_else(is.na(Station) & Negative_control==TRUE | Task=="NTC", `Sample Name`, Station)) %>% 
  mutate(., Name_plotting = str_c(Station," Bottle ", Biological_Replicate," PCR ",Technical_Replicate )) %>% 
  mutate(., Name_plotting = if_else(is.na(Name_plotting),str_c(`Sample Name`, " PCR ",Technical_Replicate),Name_plotting)) %>% 
   mutate(., Name_plotting_b = str_c(Station," Bottle ", Biological_Replicate)) %>% 
  mutate(., Name_plotting_b = if_else(is.na(Name_plotting_b),`Sample Name`,Name_plotting_b)) %>% 
  mutate(., Type = if_else(is.na(Station),"Extraction Blank",Station )) %>% 
   mutate(., Type = if_else(str_detect(Type,".NC"),"Field Blank",Type )) %>% 
  mutate(., Sample_Type = if_else(Negative_control==FALSE,"Sample","Blank")) -> pycno_sample_data_4_plotting_run1
```

### Figure sa3
```{r}
pycno_sample_data_4_plotting_run1

pycno_sample_data_4_plotting_run1$Type2 <- pycno_sample_data_4_plotting_run1$Type
pycno_sample_data_4_plotting_run1$Type2 <- as.factor(pycno_sample_data_4_plotting_run1$Type2)
levels(pycno_sample_data_4_plotting_run1$Type2)[levels(pycno_sample_data_4_plotting_run1$Type2) == "Aquarium Discharge"] <- "Aquarium\nDischarge"
levels(pycno_sample_data_4_plotting_run1$Type2)[levels(pycno_sample_data_4_plotting_run1$Type2) == "FHL Dock"] <- "FHL\nDock"
levels(pycno_sample_data_4_plotting_run1$Type2)[levels(pycno_sample_data_4_plotting_run1$Type2) == "FHL Intake Pipes"] <- "FHL\nIntake\nPipes"

pycno_sample_data_4_plotting_run1 %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., Concentration=if_else(Concentration==0,NA,Concentration) ) %>% 
  mutate(., Plate = "1")-> plate_1_data_analysis

plate_1_data_analysis %>% 
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=log10(Concentration), color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black") + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> sa3

sa3


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_run1.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```

```{r}
pycno_sample_data_4_plotting_run1 %>% 
  filter(., str_detect(Station, "Aquarium")) %>% 
  ungroup() %>% 
  dplyr::summarise(mean(Concentration), sd(Concentration))
```

### Figure sa4
```{r}

pycno_sample_data_4_plotting_run1 %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT = replace(CT, CT>40, NA)) %>% 
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=CT, color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black", limits = c(20, 40)) + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> sa4

sa4


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_Ct_run1.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```


# Load Data Pycno qPCR Plate 3

```{r}
sample_set_up_run3 <- read_excel(here("qPCR_raw_data","Pycno_test_3.xls"), sheet="Sample Setup", skip = 46)

amplification_data_run3 <- read_excel(here("qPCR_raw_data","Pycno_test_3.xls"), sheet="Amplification Data",  skip = 46)

results_data_run3<- read_excel(here("qPCR_raw_data","Pycno_test_3.xls"), sheet="Results",  skip = 46)

```

### Merge Tables
```{r}

amplification_data_run3 %>% 
  left_join(sample_set_up_run3) %>% 
  mutate(., `Sample Name` = if_else(is.na(`Sample Name`),str_c(`Target Name`,"_",Task,"_", Quantity), `Sample Name` )) %>% 
  mutate(., `Sample Name` = if_else(is.na(`Sample Name`),"Negative Control", `Sample Name` ))-> combined_qPCR_data_run3

```

Now we are ready to start analyzing the data.

First thing we want to do is check to see how our internal positive control is working.

## Moa Map
```{r}

columns = seq(from =1, to=24, by=1)
columns_t= as.tibble(columns) %>%  dplyr::rename("Column"=value)
rows = capitalize(letters[1:16])
rows_t= as.tibble(rows) %>%  dplyr::rename("Row"=value)

cross_join(columns_t, rows_t) -> full_plate


combined_qPCR_data_run3 %>% 
  filter(., `Target Name`=="MOA") %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> moa_samples_run3
  
full_plate %>% 
  left_join(moa_samples_run3) -> moa_plate_run3


moa_plate_run3 %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns ) + ylab("Row") +theme_pubr() +scale_fill_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Chevalier1"), wes_palette("Cavalcanti1"), wes_palette("Zissou1"), wes_palette("Royal1"), wes_palette("Royal2"),wes_palette("AsteroidCity1"),wes_palette("AsteroidCity2"), wes_palette("FrenchDispatch"),wes_palette("IsleofDogs1"),wes_palette("IsleofDogs2"))) +ggtitle("Moa-IPC Standards & Samples")


```



## Pycno Map
```{r}

combined_qPCR_data_run3 %>% 
  filter(., `Target Name`=="Pycno") %>% 
  mutate(., Row = str_sub(`Well Position`,start = 1L, end = 1L), 
         Column=  as.numeric(str_sub(`Well Position`,start = 2L, end = -1L ))) -> pycno_samples_run3
  
full_plate %>% 
  left_join(pycno_samples_run3) -> pycno_plate_run3


pycno_plate_run3 %>% 
  mutate(Row = factor(Row, levels = rows)) %>%  
  ggplot(aes(x=Column, y=fct_rev(Row), fill=`Sample Name` )) +geom_tile() +scale_x_continuous(breaks=columns, labels =columns ) + ylab("Row") +theme_pubr() +scale_fill_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Chevalier1"), wes_palette("Cavalcanti1"), wes_palette("Zissou1"), wes_palette("Royal1"), wes_palette("Royal2"),wes_palette("AsteroidCity1"),wes_palette("AsteroidCity2"), wes_palette("FrenchDispatch"),wes_palette("IsleofDogs1"),wes_palette("IsleofDogs2"))) +ggtitle("Pycno Standards & Samples")

```


## Moa Rn
```{r}
combined_qPCR_data_run3 %>% 
  filter(`Target Name`=="MOA" & Task=="STANDARD") -> moa_standard_run3

moa_standard_run3$Rn %>%  max() -> max_Rn_moa_run3
h_moa_run3= max_Rn_moa_run3*0.25

moa_standard_run3 %>% 
ggplot(., aes(y=Rn, x=Cycle, color=`Well Position`)) +
  geom_point(alpha=0.2) + geom_hline(yintercept =h_moa_run3) +
  geom_text(aes(0,h_moa_run3,label = round(h_moa_run3,2), vjust = -1),color="black") +theme_bw() +ggtitle("Moa Rn") +theme(legend.position = "none")


```


There is quite a bit of spread in our IPC. Will remove any samples outside normal distribution.
#### Remove Samples with Anomolous Outlier IPC values
```{r}

moa_standard_run3 %>% 
  filter(., Cycle ==40) -> moa_standard_c40_run3

moa_standard_c40_run3 %>% 
 dplyr::summarise(mean_Rn = ci(Rn, confidence = 0.99)[1], 
                      lower_ci_mpg = ci(Rn, confidence = 0.99)[2],
                      upper_ci_mpg = ci(Rn, confidence = 0.99)[3], 
                      sd_Rn = ci (Rn, confidence = 0.99)[4]) -> CI0.95_moa_run3

Sum_run3 = groupwiseMean(Rn ~ Cycle, data=moa_standard_c40_run3, conf = 0.99)

moa_standard_c40_run3 %>% 
summarise(
    Q1 = quantile(Rn, 0.25, na.rm = TRUE),
    Q3 = quantile(Rn, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR
  ) -> moa_iqr_run3

moa_standard_c40_run3 %>% 
  mutate(., type=case_when(Rn > moa_iqr_run3$upper_bound ~"Outlier High",
                           Rn < moa_iqr_run3$lower_bound ~"Outlier Low",
                           TRUE~"To Keep")) -> moa_standard_c40_run3

moa_standard_c40_run3 %>% 
  ggplot(., aes(x= Rn, y=Cycle)) +  geom_dots(layout='weave', side = "top", aes(color=type)) + stat_slabinterval(side = "bottom",aes(fill = after_stat(level)))  +scale_fill_brewer(na.translate = FALSE) +theme_pubclean()

moa_standard_c40_run3 %>% 
  filter(., type!="To Keep") -> moa_outliers_to_drop_run3
```

#### Filter Data
```{r}
combined_qPCR_data_run3 %>% 
  filter(., `Well Position` != moa_outliers_to_drop_run3$`Well Position`) -> combined_qPCR_data_cleaned_run3


results_data_run3 %>% 
    filter(., `Well Position` != moa_outliers_to_drop$`Well Position`) -> results_data_cleaned_run3
```




## Pycno Rn
```{r}
combined_qPCR_data_run3 %>% 
   filter(`Target Name`=="Pycno" & Task=="STANDARD")-> pycno_standard_run3

pycno_standard_run3$Rn %>%  max() -> max_Rn_run3
h_run3= 0.8

pycno_standard_run3 %>% 
ggplot(., aes(y=Rn, x=Cycle, color=`Sample Name`)) +geom_point(alpha=0.4) + geom_hline(yintercept =h_run3) +
  geom_text(aes(0,h_run3,label = round(h_run3,2), vjust = -1),color="black") +theme_bw() + ggtitle("Pycno Rn") +scale_colour_manual(values =c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))


```

#### Pycno Ct versus Quantity
```{r}

pycno_standard_run3 %>% 
  mutate(., Threshold = if_else(Rn > h_run3, "Above","Below")) -> pycno_standard_run3

pycno_standard_run3 %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="first")==1) %>% 
  filter(., Threshold =="Above") -> pycno_standard_above_run3

pycno_standard_run3 %>% 
  ungroup() %>% 
  filter(., !`Well Position` %in%  pycno_standard_above_run3$`Well Position`) %>% 
  group_by(`Well Position`,`Sample Name`) %>% 
  filter(., rank(Threshold, ties.method="last")==1) -> pycno_standard_below_run3

bind_rows(pycno_standard_above_run3, pycno_standard_below_run3) -> pycno_standard_threshold_run3



set_breaks = function(limits) {
     seq(limits[1], limits[2], by = 1)
}

pycno_standard_threshold_run3 %>% 
    filter(., Cycle < 40 ) %>% 
  ggplot(., aes(x=`Quantity`, y = Cycle)) +geom_count(aes(color = ..n.., size = ..n..))  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct") -> run3_ct_versus_quant


run3_ct_versus_quant
```

```{r}

# View the structure of the generated plot object
# The computed data is stored within the plot object
grob_data_run3 <- ggplot_build(run3_ct_versus_quant)$data[[3]]

# The result is a list; unlist it to get a character vector
coef_char_run3 <- unlist(grob_data_run3$coefs)


standard_curve_coefs_run3 <- enframe(coef_char_run3, name = "Metric", value = "Value") %>% 
 mutate( Metric = recode(Metric,
                             "(Intercept)" = "Intercept",
                             "x" = "Slope"))

# Extract the R^2 value from the 'r.squared' column
r_squared_value_run3 <- grob_data_run3$r.squared

```

```{r}
results_data_cleaned_run3 %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task=="STANDARD") %>% 
   mutate(., CT=as.numeric(CT)) -> pycno_results_standard_run3


 pycno_results_standard_run3 %>% 
  ggplot(., aes(x=`Quantity`, y = CT)) +geom_point()  +
    scale_x_log10("Quantity",
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))) +
  guides(color = 'legend') +
     scale_size_continuous(breaks = set_breaks)+
     scale_color_continuous(breaks = set_breaks)+
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),label.x = "right",
  label.y = "top") +theme_bw() +ylab("Ct")
 
```



```{r}
results_data_cleaned_run3 %>% 
  filter(., `Target Name`==Target_name) %>% 
  filter(., Task!="UNKNOWN") %>% 
   mutate(., CT=as.numeric(CT))  %>% 
  mutate(., Lab="NOAA PMEL OME") %>% 
  dplyr::select(Target=`Target Name`, Lab, Cq=CT, SQ=Quantity) %>% 
   mutate(., TemplateVolume =2,
         ReactionVolume= 10) %>% 
  mutate(., Cq = replace(Cq, Cq>40, NA))-> qPCR_Pycno_3_standards

write.csv(qPCR_Pycno_3_standards, file = here("LOD_LOQ/qPCR_Pycno_3_standards.csv"), row.names = F)
```

```{r}
results_data_cleaned_run3 %>% 
  filter(., `Target Name` == "Pycno") %>% 
  filter(., str_detect(`Sample Name`,"FH25"))

```

## FHL Analysis Transplanted Individual Detection

## Merge Sample Metadata
```{r}

sample_metadata<- read_excel(here("qPCR_raw_data","ElectronicArchive_CruiseSampleMetadata_20250516_FHL.xlsx"), sheet="Electronic Archive")

results_data_cleaned_run3 %>% 
   filter(.,`Target Name`=="Pycno") %>% 
  filter(., !str_detect(`Task`,"STANDARD")) %>% 
    mutate(., Quantity=if_else(is.na(Quantity), 0, Quantity)) %>% 
  mutate(., Concentration = (Quantity*10*(100/2)/ 1) %>% # Quantity is the concentration generated by the qPCR machine in copies per µL, 10 µL per reaction, 100 µL in the total extraction, 2 µL of template added, and 1 L sampled
  mutate(., Observation = if_else(Quantity ==0, "Not Detected", "Detected")) %>%
  mutate(., `Sample Name`= if_else(str_detect(`Sample Name`,"Negative Control"), "NTC",`Sample Name`)) %>% 
  ungroup() %>% 
  mutate(., Sample_Name= str_remove(`Sample Name`,".FH25")) %>% 
  left_join(sample_metadata, by=c("Sample_Name"="Sample_Name")) %>% 
  mutate(., `Sample Name` = if_else(`Sample Name`=="Ext_blank_UWFH25","Extraction Blank",`Sample Name`)) %>% 
  mutate(., Negative_control = if_else(is.na(Negative_control), TRUE, Negative_control)) %>% 
  group_by(`Sample Name`) %>% 
  mutate(Technical_Replicate = row_number(`Sample Name`)) %>% 
  mutate(., Station = if_else(is.na(Station) & Negative_control==TRUE | Task=="NTC", `Sample Name`, Station)) %>% 
  mutate(., Name_plotting = str_c(Station," Bottle ", Biological_Replicate," PCR ",Technical_Replicate )) %>% 
  mutate(., Name_plotting = if_else(is.na(Name_plotting),str_c(`Sample Name`, " PCR ",Technical_Replicate),Name_plotting)) %>% 
   mutate(., Name_plotting_b = str_c(Station," Bottle ", Biological_Replicate)) %>% 
  mutate(., Name_plotting_b = if_else(is.na(Name_plotting_b),`Sample Name`,Name_plotting_b)) %>% 
  mutate(., Type = if_else(is.na(Station),"Extraction Blank",Station )) %>% 
   mutate(., Type = if_else(str_detect(Type,".NC"),"Field Blank",Type )) %>% 
  mutate(., Sample_Type = if_else(Negative_control==FALSE,"Sample","Blank")) -> pycno_sample_data_4_plotting_run3
```

### Figure sa5
```{r}


pycno_sample_data_4_plotting_run3$Type2 <- pycno_sample_data_4_plotting_run3$Type
pycno_sample_data_4_plotting_run3$Type2 <- as.factor(pycno_sample_data_4_plotting_run3$Type2)
levels(pycno_sample_data_4_plotting_run3$Type2)[levels(pycno_sample_data_4_plotting_run3$Type2) == "Aquarium Discharge"] <- "Aquarium\nDischarge"
levels(pycno_sample_data_4_plotting_run3$Type2)[levels(pycno_sample_data_4_plotting_run3$Type2) == "FHL Dock"] <- "FHL\nDock"
levels(pycno_sample_data_4_plotting_run3$Type2)[levels(pycno_sample_data_4_plotting_run3$Type2) == "FHL Intake Pipes"] <- "FHL\nIntake\nPipes"

pycno_sample_data_4_plotting_run3 %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., Concentration=if_else(Concentration==0,NA,Concentration) ) %>% 
  mutate(., Plate = "3")-> plate_3_data_analysis
  
plate_3_data_analysis %>%   
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=log10(Concentration), color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black") + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> sa5

sa5


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_run3.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```

### Figure sa6
```{r}

pycno_sample_data_4_plotting_run3 %>% 
  filter(., str_detect(Field_collection_method,"Sterivex")) %>% 
  filter(., Sample_Type !="Blank") %>% 
       mutate(across(Sample_Type, ~factor(., levels=c("Sample","Blank")))) %>% 
  group_by(Name_plotting_b) %>% 
  mutate(., Replicate = row_number()) %>% 
  ungroup() %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT = replace(CT, CT>40, NA)) %>% 
  ggplot(., aes(x=Name_plotting_b, y=Replicate, fill=CT, color="")) + geom_tile() +theme_pubr()+theme(axis.text.x = element_text(angle = -65, vjust = 1, hjust=-0.05))  +facet_wrap(.~Type2, nrow=1, scales="free_x")  + xlab("Bottle & Sterivex\nFilter Replicate") + ylab("Technical qPCR Replicate") +theme(axis.text.x=element_blank(),axis.text.y=element_blank(), axis.ticks.x=element_blank(), axis.title = element_text(size = 16, face = "bold"), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) + scale_fill_viridis_c(option="magma", begin = 0.2, end=1, na.value = "black", limits = c(20, 40)) + scale_colour_manual(values=NA) +   
     guides(colour=guide_legend("No Detection", override.aes=list(colour="black"))) +
  labs(fill = "log10(Copies per L)") -> sa6

sa6


ggsave(filename = here("Figures","FHL_mesocosm_field_trial_Ct_run3.png"),
       width=10,
       height = 4,
       dpi = 400,
      units = c("in"))

```

# Combine Standard Plates
```{r}

plate1_standards <- read.delim(file = here("LOD_LOQ/qPCR_Pycno_1_standards.csv"), header=T, sep=",") %>% mutate(., Plate = 1)

plate3_standards <- read.delim(file = here("LOD_LOQ/qPCR_Pycno_3_standards.csv"), header=T, sep=",") %>% mutate(., Plate = 3)

plate5_standards <- read.delim(file = here("LOD_LOQ/qPCR_Pycno_5_v3_standards.csv"), header=T, sep=",") %>% mutate(., Plate = 5)

rbind(plate1_standards, plate3_standards, plate5_standards) -> plates_135_combined

write.csv(plates_135_combined, file = here("LOD_LOQ/plates_135_combined.csv"), row.names = F)

```

```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  ggplot(aes(x= log10(SQ), y=Cq, color= Plate)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")))
  


```

```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  ggplot(aes(x= log10(SQ), y=Cq, color= Plate)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +facet_wrap(~Plate)
  

```

```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  ggplot(aes(x= log10(SQ), y=Cq, color= Plate)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +facet_wrap(~Plate)
  

```




```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  ggplot(aes(x= log10(SQ), y=Cq)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) 
  

```

```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  filter(, Plate !="1") %>% 
  ggplot(aes(x= log10(SQ), y=Cq)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) 
  

```

```{r}

plates_135_combined %>%
  mutate(., Plate= as.character(Plate)) %>% 
  mutate(., SQ=SQ*TemplateVolume) %>% 
  filter(, Plate !="1") %>% 
  ggplot(aes(x= log10(SQ), y=Cq, color=Plate)) + geom_point() + stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) 
  

```
```{r}


bind_rows(plate_5_data_analysis,plate_3_data_analysis,plate_1_data_analysis) -> combo_plate_analysis

combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  ggplot(., aes(x=Name_plotting, y= CT, fill= Plate)) +geom_col(stat = "identity",position="dodge") +facet_wrap(~Name_plotting_b, scales="free_x")
```

```{r}

combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  ggplot(., aes(x=Name_plotting_b, y= CT, color= Plate)) +geom_count() +facet_wrap(~Type2, scales="free_x")


```

```{r}

combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  filter(., CT >0) %>% 
  ggplot(., aes(x=Name_plotting_b, y= CT, color= Plate)) +geom_boxplot() +facet_wrap(~Type2, scales="free_x")


```

```{r}
combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  dplyr::group_by(Plate, Type, Name_plotting_b) %>% 
  filter(., CT > 0) %>% 
  dplyr::summarise(mean_CT= mean(CT), sd_CT=sd(CT), n=n())
```

```{r}

combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  dplyr::group_by(Plate, Type, Name_plotting_b) %>% 
  filter(., CT > 0) %>% 
  dplyr::summarise(mean_CT= mean(CT), sd_CT=sd(CT), n=n())
```

```{r}
combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  dplyr::group_by(Plate, Type) %>% 
  filter(., CT > 0) %>% 
  dplyr::summarise(mean_CT= mean(CT), sd_CT=sd(CT), n=n())

```

```{r}
combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  dplyr::group_by(Plate, Type) %>% 
  filter(., CT < 1) %>% 
  dplyr::summarise(n=n())

```
```{r}
combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  filter(., CT > 0)  -> fhl_pos_CT

model_aov_fhl_pos_CT <- aov(CT ~ Plate + Type + Plate*Type, data=fhl_pos_CT)

summary(model_aov_fhl_pos_CT)

tukey_hsd(model_aov_fhl_pos_CT)
```

```{r}

model_aov_fhl_pos_CT_bottle <- aov(CT ~ Plate + Name_plotting_b + Plate*Name_plotting_b, data=fhl_pos_CT)

summary(model_aov_fhl_pos_CT_bottle)

tukey_hsd(model_aov_fhl_pos_CT_bottle)
```


```{r}
combo_plate_analysis %>% 
  mutate(., CT=as.numeric(CT)) %>% 
  mutate(., CT= replace_na(CT, 0)) %>% 
  filter(., CT > 0) %>% 
  filter(., Type =="FHL Dock" ) -> fhl_dock_CT

model_aov_fhl_dock <- aov(CT ~ Plate, data=fhl_dock_CT)

summary(model_aov_fhl_dock)

tukey_hsd(model_aov_fhl_dock)
```

# Figure 5

```{r}
pycno_sample_data_4_plotting  %>% 
  mutate(., Plate="Pycno Run 5") %>% 
  dplyr::select(Plate, `Sample Name`, CT, Quantity, `Quantity Mean`, `Quantity SD`, Concentration, Biological_Replicate, Technical_Replicate, Name_plotting, Name_plotting_b, Station)-> LoD_plate3_analysis
pycno_sample_data_4_plotting_run3 %>% 
  mutate(., Plate="Pycno Run 3") %>% 
  dplyr::select(Plate, `Sample Name`, CT, Quantity, `Quantity Mean`, `Quantity SD`, Concentration, Biological_Replicate, Technical_Replicate, Name_plotting, Name_plotting_b, Station)-> LoD_plate2_analysis
pycno_sample_data_4_plotting_run1 %>% 
  mutate(., Plate="Pycno Run 1") %>% 
  dplyr::select(Plate, `Sample Name`, CT, Quantity, `Quantity Mean`, `Quantity SD`, Concentration, Biological_Replicate, Technical_Replicate, Name_plotting, Name_plotting_b, Station)-> LoD_plate1_analysis

colnames(LoD_plate3_analysis)
colnames(LoD_plate2_analysis)
colnames(LoD_plate1_analysis)

bind_rows(LoD_plate3_analysis, LoD_plate2_analysis, LoD_plate1_analysis) -> LoD_plate_analysis_combo

loblodloq_df <- read.delim(, file=here("LOD_LOQ","final_lob_lod_loq_table.csv"), sep=",") %>% 
  filter(., Type=="qPCR") %>% pivot_wider(names_from = LoD_method, values_from = c(LOB,LOD,LOQ)) %>% 
  mutate(., LOB_eLowQuant = if_else(is.na(LOB_eLowQuant),0, LOB_eLowQuant),
         LOD_eLowQuant = if_else(is.na(LOD_eLowQuant),0.2134723, LOD_eLowQuant),
         LOQ_eLowQuant = if_else(is.na(LOQ_eLowQuant),0.8078273, LOQ_eLowQuant))

wes_palette("AsteroidCity1") -> AC

LoD_plate_analysis_combo %>% 
  left_join(loblodloq_df) %>% 
  filter(., Station %in% c("FHL Dock","FHL Intake Pipes")) %>% 
  mutate(., Quantity=Quantity*2) %>% 
  ggplot(aes(y=Quantity, x= Station, color=Plate, shape =Plate)) + geom_jitter(size=5, alpha=0.4) + theme_pubclean() + xlab("Copies per Reaction") + facet_wrap(~ Plate, scales = "free_y") +
  scale_color_manual(values =c(AC[2],AC[4],AC[5])) + 
geom_hline(aes(yintercept=LOD_eLowQuant),color=AC[1]) +
  geom_text(aes(y = LOD_eLowQuant, x = Inf, label = paste0("eLowQuant LoD = ", round(LOD_eLowQuant, 2))), 
             vjust = -4.5, hjust=1.4, color = AC[1])+ geom_hline(aes(yintercept=LOD_Klymus), color=AC[3])  +
  geom_text(aes(y = LOD_Klymus, x = Inf, label = paste0("Klymus et al. 2020 LoD = ", round(LOD_Klymus, 1.4))), vjust = 2, hjust=1.2, color = AC[3]) +theme(legend.position = "none")+theme(axis.title = element_text(size = 16, face = "bold"), axis.text = element_text(size = 16), strip.text = element_text(size = 16, face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold") ) 


ggsave(filename = here("Figures","Figure_5_field_trial_LoD.png"),
       width=10,
       height = 6,
       dpi = 400,
      units = c("in"))
```




