---
title: "FAIRe_Stanford_20260915"
author: "Zachary Gold"
date: "2026-09-18"
output: html_document
---

---
title: "FAIRe_Hakai_conversion_20260917"
author: "Zachary Gold"
date: "2026-09-18"
output: html_document
---

```{r}
library(tidyverse)
library(rmarkdown)
library(here)
library(readxl)
```

# Read in Saple Metadata
```{r}
Stanford_22Jan2026_SonomaFieldSamples <- read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_22Jan2026_SonomaFieldSamples.csv") %>% 
  mutate(., Plate="SonomaFieldSamples")

Stanford_1Feb2026_LarvalSamples <- read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_1Feb2026_LarvalSamples.csv") %>% 
  mutate(., Plate="LarvalSamples")

Stanford_6Feb2026_LarvalSamplesDiluted <-read.csv("raw_data/ddPCR_raw_data/Stanford/Stanford_6Feb2026_LarvalSamplesDiluted.csv") %>% 
  mutate(., Plate="LarvalSamplesDiluted")

sample_metadata<- bind_rows(Stanford_22Jan2026_SonomaFieldSamples, Stanford_1Feb2026_LarvalSamples, Stanford_6Feb2026_LarvalSamplesDiluted)
```

# Read in FAIRe BeBOP
```{r}

# Read all lines and find the YAML boundaries
raw_lines_SR_collect_bebop <- read_lines(here("FAIRe/PMEL/NOAA-PMEL-OME_eDNA_Collection_Protocol_SmithRoot_PeristalticPump.md"))
bounds_SR_collect_bebop <- which(raw_lines_SR_collect_bebop == "---")

# Extract the header lines into a tibble
PMEL_SR_collect_bebop_list <- tibble(raw_text = raw_lines_SR_collect_bebop[(bounds_SR_collect_bebop[1] + 1):(bounds_SR_collect_bebop[2] - 1)]) |> 
  # Remove empty lines
  filter(trimws(raw_text) != "") |> 
  # Create a category column by detecting the '#' symbol and removing it
  mutate(
    category = if_else(str_starts(raw_text, "#"), str_remove(raw_text, "^#\\s*"), NA_character_)
  ) |> 
  # Fill the category names down to the subsequent rows
  fill(category, .direction = "down") |> 
  # Drop the original comment rows now that the category is assigned
  filter(!str_starts(raw_text, "#")) |> 
  # Split the text into term and value at the first colon
  # extra = "merge" ensures we don't accidentally split values that contain colons
  separate(raw_text, into = c("term", "value"), sep = ":\\s*", extra = "merge") |> 
  # Clean up any rogue whitespace
  mutate(across(everything(), str_trim))

# Read all lines and find the YAML boundaries
raw_lines_Stx_collect_bebop <- read_lines(here("FAIRe/PMEL/NOAA-PMEL-OME_eDNA_Collection_Protocol_Sterivex_PeristalticPump.md"))
bounds_Stx_collect_bebop <- which(raw_lines_Stx_collect_bebop == "---")

# Extract the header lines into a tibble

PMEL_Stx_collect_bebop_list <- tibble(raw_text = raw_lines_Stx_collect_bebop[(bounds_Stx_collect_bebop[1] + 1):(bounds_Stx_collect_bebop[2] - 1)]) |> 
  # Remove empty lines
  filter(trimws(raw_text) != "") |> 
  # Create a category column by detecting the '#' symbol and removing it
  mutate(
    category = if_else(str_starts(raw_text, "#"), str_remove(raw_text, "^#\\s*"), NA_character_)
  ) |> 
  # Fill the category names down to the subsequent rows
  fill(category, .direction = "down") |> 
  # Drop the original comment rows now that the category is assigned
  filter(!str_starts(raw_text, "#")) |> 
  # Split the text into term and value at the first colon
  # extra = "merge" ensures we don't accidentally split values that contain colons
  separate(raw_text, into = c("term", "value"), sep = ":\\s*", extra = "merge") |> 
  # Clean up any rogue whitespace
  mutate(across(everything(), str_trim))




# Read all lines and find the YAML boundaries
raw_lines_Stx_extr_bebop <- read_lines(here("FAIRe/PMEL/NOAA-PMEL-OME_eDNA_Sterivex_Extraction_Protocol_Centrifuge.md"))
bounds_Stx_extr_bebop <- which(raw_lines_Stx_extr_bebop == "---")

# Extract the header lines into a tibble
PMEL_Stx_extr_bebop_list <- tibble(raw_text = raw_lines_Stx_extr_bebop[(bounds_Stx_collect_bebop[1] + 1):(bounds_Stx_collect_bebop[2] - 1)]) |> 
  # Remove empty lines
  filter(trimws(raw_text) != "") |> 
  # Create a category column by detecting the '#' symbol and removing it
  mutate(
    category = if_else(str_starts(raw_text, "#"), str_remove(raw_text, "^#\\s*"), NA_character_)
  ) |> 
  # Fill the category names down to the subsequent rows
  fill(category, .direction = "down") |> 
  # Drop the original comment rows now that the category is assigned
  filter(!str_starts(raw_text, "#")) |> 
  # Split the text into term and value at the first colon
  # extra = "merge" ensures we don't accidentally split values that contain colons
  separate(raw_text, into = c("term", "value"), sep = ":\\s*", extra = "merge") |> 
  # Clean up any rogue whitespace
  mutate(across(everything(), str_trim))




# Read all lines and find the YAML boundaries
raw_lines_SR_extr_bebop <- read_lines(here("FAIRe/PMEL/NOAA-PMEL-OME_eDNA_DiscFilters_Extraction_Protocol_Centrifuge.md"))
bounds_SR_extr_bebop <- which(raw_lines_SR_extr_bebop == "---")

# Extract the header lines into a tibble

PMEL_SR_extr_bebop_list <- tibble(raw_text = raw_lines_SR_extr_bebop[(bounds_SR_extr_bebop[1] + 1):(bounds_SR_extr_bebop[2] - 1)]) |> 
  # Remove empty lines
  filter(trimws(raw_text) != "") |> 
  # Create a category column by detecting the '#' symbol and removing it
  mutate(
    category = if_else(str_starts(raw_text, "#"), str_remove(raw_text, "^#\\s*"), NA_character_)
  ) |> 
  # Fill the category names down to the subsequent rows
  fill(category, .direction = "down") |> 
  # Drop the original comment rows now that the category is assigned
  filter(!str_starts(raw_text, "#")) |> 
  # Split the text into term and value at the first colon
  # extra = "merge" ensures we don't accidentally split values that contain colons
  separate(raw_text, into = c("term", "value"), sep = ":\\s*", extra = "merge") |> 
  # Clean up any rogue whitespace
  mutate(across(everything(), str_trim))

# Read all lines and find the YAML boundaries
raw_lines_qPCR_bebop <- read_lines(here("FAIRe/PMEL/NOAA-PMEL-OME_Pycno_nad5_v1_qPCR_Protocol_BeBOP.md"))
bounds_qPCR_bebop <- which(raw_lines_qPCR_bebop == "---")

# Extract the header lines into a tibble

PMEL_qPCR_bebop_list <- tibble(raw_text = raw_lines_qPCR_bebop[(bounds_qPCR_bebop[1] + 1):(bounds_qPCR_bebop[2] - 1)]) |> 
  # Remove empty lines
  filter(trimws(raw_text) != "") |> 
  # Create a category column by detecting the '#' symbol and removing it
  mutate(
    category = if_else(str_starts(raw_text, "#"), str_remove(raw_text, "^#\\s*"), NA_character_)
  ) |> 
  # Fill the category names down to the subsequent rows
  fill(category, .direction = "down") |> 
  # Drop the original comment rows now that the category is assigned
  filter(!str_starts(raw_text, "#")) |> 
  # Split the text into term and value at the first colon
  # extra = "merge" ensures we don't accidentally split values that contain colons
  separate(raw_text, into = c("term", "value"), sep = ":\\s*", extra = "merge") |> 
  # Clean up any rogue whitespace
  mutate(across(everything(), str_trim))

```

```{r}
bind_rows(PMEL_SR_collect_bebop_list,PMEL_Stx_collect_bebop_list,PMEL_Stx_extr_bebop_list,PMEL_SR_extr_bebop_list,PMEL_qPCR_bebop_list) -> bebop_total

bebop_total %>% 
  filter(., category =="MIOP terms") %>% 
  distinct() -> miop_terms

miop_terms %>% 
  group_by(term) %>% 
  summarise(value = paste(value, collapse = ";"), .groups = "drop") %>% 
  mutate(., value = if_else(term=="creator","Han Weinrich, Zachary Gold, Shannon Brown, Sean McAllister",value)) %>% mutate(., `term:value`=str_c(term,":",value)) %>% ungroup() %>% dplyr::select(`term:value`) -> miop_unique

as.list(miop_unique) -> miop_unique_list

miop_unique_values <- paste(miop_unique_list$`term:value`, collapse = "|")
```

# Read in FAIRe Checklist information
```{r}
FAIRe_projectMetadata_checklist <- read_xlsx(here("FAIRe/FAIRe_checklist_v1.0.2_FULLtemplate.xlsx"), sheet="projectMetadata")

FAIRe_sampleMetadata_checklist <- read_xlsx(here("FAIRe/FAIRe_checklist_v1.0.2_FULLtemplate.xlsx"), sheet="sampleMetadata") 

FAIRe_sampleMetadata_checklist <- unlist(FAIRe_sampleMetadata_checklist[2, ], use.names = FALSE) %>% as_tibble_col(column_name = "term") 

FAIRe_stdData_checklist <- read_xlsx(here("FAIRe/FAIRe_checklist_v1.0.2_FULLtemplate.xlsx"), sheet="stdData") 

FAIRe_stdData_checklist <- unlist(FAIRe_stdData_checklist[2, ], use.names = FALSE) %>% as_tibble_col(column_name = "term") 

FAIRe_ampData_checklist <- read_xlsx(here("FAIRe/FAIRe_checklist_v1.0.2_FULLtemplate.xlsx"), sheet="ampData") 

FAIRe_ampData_checklist <- unlist(FAIRe_ampData_checklist[2, ], use.names = FALSE) %>% as_tibble_col(column_name = "term")

FAIRe_eLowQuantData_checklist <- read_xlsx(here("FAIRe/FAIRe_checklist_v1.0.2_FULLtemplate.xlsx"), sheet="eLowQuantData") 

FAIRe_eLowQuantData_checklist <- unlist(FAIRe_eLowQuantData_checklist[2, ], use.names = FALSE) %>% as_tibble_col(column_name = "term") 


```

### lod
```{r}

loblodloq_df <- read.delim(, file=here("LOD_LOQ","final_lob_lod_loq_table.csv"), sep=",") %>% 
  filter(., assay_method=="ddPCR" & Lab=="Stanford" & Type=="1 Droplet") %>% pivot_wider(names_from = LoD_method, values_from = c(LOB,LOD,LOQ)) 
```

# Project Metadata

## Mapping of Values
```{r}

sample_metadata %>% 
   mutate(recordedBy ="Meghan M. Shea") %>% 
  mutate(., recordedByID= "https://orcid.org/0000-0002-7419-6654",
         project_contact = "zachary.gold@noaa.gov",
         institution= "Stanford University",
         institutionID= NA,
         project_name = "Molecular Star Gazing: Development and Validation of an Environmental DNA Assay for the Imperiled Sunflower Sea Star (Pycnopodia helianthoides)",
         project_id = "Pycnopodia_eDNA_assay_validation_MS",
         parent_project_id = "NOAA_PMEL_OME",
         study_factor = miop_unique_values,
         assay_type = "targeted",
         sterilise_method = "https://doi.org/10.5281/zenodo.16945434",
         neg_cont_0_1 = 1,
         pos_cont_0_1 = 1,
         checkls_ver = "1.0.2",
         mod_date = "2026-09-17",
         license = "CC0 1.0 Universal",
         rightsHolder ="NOAA PMEL",
         accessRights="Universal",
         informationWithheld=NA,
         dataGeneralizations=NA,
         bibliographicCitation="https://doi.org/10.5281/zenodo.20076925",
         associated_resource="https://doi.org/10.64898/2026.05.07.723600",
         code_repo="https://github.com/marinednadude/Pycnopodia_eDNA_assay_validation_MS",
         biological_rep = NA,
         pcr_0_1 = 1,
         inhibition_check_0_1=0,
         inhibition_check=NA,
         thermocycler = "AutoDG Automated Droplet Generator (Bio-Rad, Hercules, CA), a C1000 Touch Thermal Cycler (Bio-Rad, Hercules, CA) and a QX200 Droplet Digital PCR System (Bio-Rad, Hercules, CA)",
         assay_name=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="assay_name")],
         assay_validation=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="assay_validation")],
         targetTaxonomicAssay=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="targetTaxonomicAssay")],
        targetTaxonomicScope="",
        target_gene=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="target_gene")],
        target_subfragment="",
        target_subfragment="",
        ampliconSize=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="ampliconSize")],
        pcr_primer_forward=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_forward")],
        pcr_primer_reverse=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_reverse")],
        pcr_primer_name_forward=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_name_forward")],
        pcr_primer_name_reverse=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_name_reverse")],
        pcr_primer_reference_forward="https://doi.org/10.64898/2026.05.07.723600",
        pcr_primer_reference_reverse="https://doi.org/10.64898/2026.05.07.723600",
        pcr_primer_vol_forward=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_vol_forward")],
        pcr_primer_vol_reverse=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_vol_reverse")],                                                                              
        pcr_primer_conc_forward=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_conc_forward")],
        pcr_primer_conc_reverse=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_primer_conc_reverse")],
        probeReporter=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="probeReporter")],
        probeQuencher=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="probeQuencher")],
        probeQuencher=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="probeQuencher")],
        probe_seq=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="probe_seq")],
        probe_ref="https://doi.org/10.64898/2026.05.07.723600",
        probe_conc=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="probe_conc")],
        commercial_mm=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="commercial_mm")],
        custom_mm=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="custom_mm")],
        block_seq=NA,
        block_ref=NA,
        block_taxa=NA,
        amplificationReactionVolume=20,
        pcr_dna_vol=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_dna_vol")],
        pcr_rep=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="pcr_rep")],
        nucl_acid_amp="https://doi.org/10.64898/2026.05.07.723600",
        pcr_cond="denaturation:95_10;annealing:95_0.5;anealing:57_1;45;Extension:98_10",
        annealingTemp=57,
        pcr_cycles=35,
        pcr_analysis_software="R | Bio-Rad QX Insight ddPCR Analysis Software",
        pcr_method_additional="AutoDG Automated Droplet Generator (Bio-Rad, Hercules, CA), a C1000 Touch Thermal Cycler (Bio-Rad, Hercules, CA) and a QX200 Droplet Digital PCR System (Bio-Rad, Hercules, CA) (Shea et al., 2026) https://doi.org/10.5281/zenodo.20057581",
         amp_vis_method="ddPCR",
        detection_criteria=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="detection_criteria")],
        lod_method=paste0(PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="lod_method")],"|"," https://doi.org/10.1002/edn3.220"),
        pcr_assay_lod_techreps=8,
        pcr_assay_lod=3.25,
        pcr_assay_lod_unit="copies/reaction",
        pcr_assay_lod_UL=4.6762058613743,
        pcr_assay_lod_LL=1.83303099951927,
        loq_method=paste0(PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="loq_method")],"|"," https://doi.org/10.1002/edn3.220"),
        pcr_assay_loq_techreps=8,
        pcr_assay_loq=11,
        pcr_assay_loq_unit="copies/reaction",
        pcr_assay_loq_UL=NA,
        pcr_assay_loq_LL=NA,
        elowquant_lod_method	=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="elowquant_lod_method")],
                                                         elowquant_pcr_assay_lod_techreps	=8,
        elowquant_pcr_assay_lod	=0.132715432406258,
        elowquant_pcr_assay_lod_unit	="copies/reaction",
        elowquant_pcr_assay_lod_UL=36.4466534194249,
        elowquant_pcr_assay_lod_LL	=0,
        elowquant_loq_method			=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="elowquant_loq_method")],
        elowquant_pcr_assay_loq_techreps	=8,
        elowquant_pcr_assay_loq	=0.502225198843848,
        elowquant_pcr_assay_loq_unit	="copies/reaction",
        elowquant_pcr_assay_loq_UL	=140.706601978664,
        elowquant_pcr_assay_loq_LL=-0.0237901213192615,
        std_type=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="std_type")],
        std_source=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="std_source")],
        std_seq=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="std_seq")],
        thresholdQuantificationCycle=35,
        baselineValue=10,
        automaticThresholdQuantificationCycle=0,
        automaticBaselineValue=0,
        targeted_detection_method_additional="",
        estimatedNumberOfCopies_method=PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="estimatedNumberOfCopies_method")]) -> sample_projectMetadata_pycno

```


## Check for Completeness
```{r}
PMEL_qPCR_bebop_list %>% 
  filter(., category=="FAIRe terms") -> bebop_faire_qPCR_terms

setdiff(bebop_faire_qPCR_terms$term,colnames(sample_projectMetadata_pycno))

FAIRe_projectMetadata_checklist %>% 
  filter(., section %in% c("Project","Data management","Sample collection","PCR","Targeted assay detection")) -> faire_projectMetadata_checklist_targeted

setdiff(faire_projectMetadata_checklist_targeted$term_name,colnames(sample_projectMetadata_pycno))


```

## Make Dataframe and export TSV

```{r}

sample_projectMetadata_pycno %>% 
  ungroup() %>% 
  slice(1) %>% 
    mutate(across(everything(), as.character)) %>% 
  pivot_longer(everything(), names_to = "term_name", values_to = "project_level") -> projectMetadata_pycno_long

faire_projectMetadata_checklist_targeted %>% 
  add_row(requirement_level_code="O",	`section`="Targeted assay detection",	term_name = "estimatedNumberOfCopies_method", project_level = NA) %>%  # add in an older FAIRe term that we used
dplyr::select(-project_level) %>% 
  left_join(projectMetadata_pycno_long) -> projectMetadata_pycno

write.table(projectMetadata_pycno, file=here("FAIRe","Stanford","Stanford_projectMetadata_pycno.tsv"),sep = "\t", row.names = FALSE, quote = FALSE)

```

# Sample Metadata

```{r}
sample_projectMetadata_pycno

sample_projectMetadata_pycno %>% 
  group_by(`Sample.description.1`) %>% 
  mutate(., group_idx=row_number()) %>% 
  ungroup() %>% 
   mutate(samp_name = paste0(Sample.description.1,"-",group_idx)) %>% 
    ungroup() %>% 
   mutate(., samp_category= case_when(SampleType=="PosCtrl" ~ "positive control",
                                      SampleType=="NTC" ~ "negative control",
          TRUE ~ "sample")) %>% 
  mutate(., neg_cont_type = case_when(samp_category=="negative control" ~"PCR negative",
                                      TRUE ~"")) %>%
  mutate(., pos_cont_type = case_when(SampleType=="PosCtrl" ~"synthetic DNA",
                                      TRUE ~"")) %>%
  mutate(., materialSampleID=str_c("Stanford-Pycno-2025","_",Sample.description.1)) -> Stanford_sampleMetadata
  
 Stanford_sampleMetadata %>% 
  group_by(Plate) %>% 
  summarise(rel_cont_id = paste(samp_name, collapse = "|"), .groups = "drop") -> associated_samples
```

```{r}

Stanford_sampleMetadata %>% 
  mutate(., date = case_when(Sample.description.1=="Wasting-0" ~ "2025-08-20",
                             Sample.description.1=="Wasting-50" ~ "2025-08-21",
                             Sample.description.1=="Healthy-0" ~ "2025-08-17",
                             Sample.description.1=="Healthy-50" ~ "2025-08-18",
                             Sample.description.1=="Healthy-100" ~ "2025-08-19",
                             TRUE ~"2026-01-23")) %>% 
  dplyr::select(samp_name,date) %>% 
  filter(., !is.na(date)) %>% 
  mutate(Collection_Date_local = as.POSIXct(date)) %>% 
    mutate(eventDate = format(Collection_Date_local, "%Y-%m-%dT%H:%M:%S")) %>%
  dplyr::select(samp_name,eventDate,verbatimEventDate=Collection_Date_local)-> event_date_info


Stanford_sampleMetadata %>% 
  mutate(., env_medium = "sea water [ENVO:00002149]",
         habitat_natural_artificial_0_1 = case_when(str_detect(Sample.description.1,"Wasting")| str_detect(Sample.description.1,"Healthy") ~0,
                                                                                                           TRUE~1)) %>% 
       dplyr::select(samp_name,env_medium,habitat_natural_artificial_0_1) -> env_medium_info



Stanford_sampleMetadata %>% 
    mutate(., sample_derived_from = materialSampleID,
           sample_composed_of =materialSampleID) %>% 
    left_join(associated_samples) %>% 
  mutate(., decimalLatitude = case_when(str_detect(Sample.description.1,"Wasting")| str_detect(Sample.description.1,"Healthy") ~38.724848,
TRUE~36.805258),
decimalLongitude = case_when(str_detect(Sample.description.1,"Wasting")| str_detect(Sample.description.1,"Healthy") ~-123.482960,
TRUE~-121.785913)) %>% 
    mutate(., geo_loc_name = "North East Pacific Ocean [GAZ:00013765]") %>% 
   left_join(event_date_info) %>% 
    mutate(., env_broad_scale = "marine biome [ENVO:00000447]",
          env_local_scale ="marine benthic biome [ENVO:01000024]") %>% 
    left_join(env_medium_info) %>% 
    mutate(., samp_collect_method = "https://dx.doi.org/10.17504/protocols.io.bp2l69y7klqe/v2",
           samp_collect_device="bottle",
           samp_size = Volume.Filtered,
           samp_size_unit = "mL",
           samp_store_temp	=-20,
           samp_store_sol=NA,
           samp_store_dur = "P6M",
           samp_store_loc="Boehm Laboratory, Stanford University",
           dna_store_loc="Boehm Laboratory, Stanford University",
           samp_store_method_additional=NA) %>% 
         mutate(.,  samp_mat_process="Seawater samples were filtered through a 0.22 µm PES Sterivex sterile filter capsule to capture eDNA and DNA-containing particulates using graity filtration.") %>% 
           mutate(., filter_passive_active_0_1= 1) %>% 
           mutate(., stationed_sample_dur = NA,
           pump_flow_rate = 0.05,
           pump_flow_rate_unit="L/min",
           prefilter_material=NA,
           size_frac_low = NA,
           size_frac=  0.2,
           filter_diameter= NA,
           filter_surface_area = 1000,
           filter_material = "polyvinylidene fluoride",
           filter_name = "Sterivex SVGV010RS",
           precip_chem_prep=NA,
           precip_force_prep=NA,
           precip_time_prep=NA,
           precip_temp_prep=NA,
           prepped_samp_store_temp=NA,
           prepped_samp_store_sol =NA,
           prepped_samp_store_dur = NA,
           prep_method_additional = NA,
           date_ext =  NA,
           samp_vol_we_dna_ext = Volume.Filtered,
           samp_vol_we_dna_ext_unit = "mL",
           nucl_acid_ext_lysis = "thermal | chemical",
           nucl_acid_ext_sep = "column-based",
           nucl_acid_ext = "https://dx.doi.org/10.17504/protocols.io.ewov1qyyygr2/v1",
           nucl_acid_ext_kit = "Qiagen Blood and Tissue Kit",
           nucl_acid_ext_modify ="Modification following Spens et al. 2017. 720 µL ATL Buffer and 80 µL Proteinase K are added to the Sterivex filters. Add Buffer AL and ice-cold 100% EtOH in equal volume to recovered volume after incubation. All other methods follow the manufacturers' instructions.",
           dna_cleanup_0_1=0,
           dna_cleanup_method=NA,
           concentration = NA,
           concentration_unit = "ng/µL",
           concentration_method = "Qubit 4.0 Fluorometer (Invitrogen)",
           assay_name = PMEL_qPCR_bebop_list$value[which(PMEL_qPCR_bebop_list$term=="assay_name")],
           detected_notDetected = if_else(Copies.20µLWell >0, "Detected", "Not Detected")) -> pycno_sampleMetadata
 
	
pycno_sampleMetadata
		

```

## Check for Completeness
```{r}
bebop_total %>% 
  filter(., category=="FAIRe terms") -> bebop_faire_total_terms

setdiff(bebop_total$term,colnames(pycno_sampleMetadata))


setdiff(FAIRe_sampleMetadata_checklist$term,colnames(pycno_sampleMetadata))


```
## Make Dataframe and export TSV

```{r}

pycno_sampleMetadata %>% 
  dplyr::select(any_of(FAIRe_sampleMetadata_checklist$term)) -> sampleMetadata_pycno

write.table(sampleMetadata_pycno, file=here("FAIRe","Stanford","Stanford_sampleMetadata_pycno.tsv"),sep = "\t", row.names = FALSE, quote = FALSE)

```

# AmpData

```{r}

pycno_sampleMetadata %>% 
mutate(., estimatedNumberOfCopies=Copies.20µLWell,
       pcr_plate_id=Plate,
        quantificationCycle=NA) %>% 
rename(., technical_rep_id= group_idx) %>% 
mutate(., estimatedNumberOfCopies_method= "Ct/Cq with standard curve",
       estimatedNumberOfCopies_unit="copies/reaction" ) -> pycno_ampData 

```


## Check for Completeness
```{r}
bebop_total %>% 
  filter(., category=="FAIRe terms") -> bebop_faire_total_terms

setdiff(bebop_total$term,colnames(pycno_sampleMetadata))

setdiff(FAIRe_ampData_checklist$term,colnames(pycno_ampData))

```
## Make Dataframe and export TSV

```{r}

pycno_ampData %>% 
  dplyr::select(any_of(FAIRe_ampData_checklist$term)) -> ampData_pycno

write.table(ampData_pycno, file=here("FAIRe","Stanford","Stanford_ampData_pycno.tsv"),sep = "\t", row.names = FALSE, quote = FALSE)

```

# eLowQuant LoD Filtered
```{r}

ampData_pycno %>% 
  mutate(., Lab="Stanford") %>% 
  left_join(loblodloq_df, by=c("Lab"="Lab")) %>% 
  mutate(., LoDLoQLevel = case_when(estimatedNumberOfCopies > LOQ_Klymus ~ "Above LOQ_Klymus & LOQ_eLowQuant",
         estimatedNumberOfCopies > LOD_Klymus ~ "Above LOD_Klymus & LOQ_eLowQuant",
         estimatedNumberOfCopies > LOQ_eLowQuant ~ "Above LOQ_eLowQuant",
              estimatedNumberOfCopies > LOD_eLowQuant ~ "Above LOD_eLowQuant",                      
         estimatedNumberOfCopies > LOB_eLowQuant ~ "Above LOB_eLowQuant",
                TRUE ~ "No Amplification" )) %>% 
  mutate(., estimatedNumberOfCopies = case_when(estimatedNumberOfCopies > LOD_eLowQuant ~ estimatedNumberOfCopies,
                         estimatedNumberOfCopies < LOD_eLowQuant ~ 0)) -> pycno_eLowQuantData
  

pycno_eLowQuantData %>% 
  dplyr::select(any_of(c(FAIRe_ampData_checklist$term,"LoDLoQLevel"))) -> eLowQuantData_pycno

write.table(eLowQuantData_pycno, file=here("FAIRe","Stanford","Stanford_eLowQuantData_pycno.tsv"),sep = "\t", row.names = FALSE, quote = FALSE)

```


