---
title: 'Lesperance eLowQuant LoD and LoQ: Hakai ddPCR for 1 droplet'
author: "Zachary Gold"
date: "2026-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")

knitr::opts_chunk$set(
  message = FALSE,  # Hides diagnostic messages
  warning = FALSE   # Hides warning messages
  
)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Packages used
packages = c("tidyverse", 
             "dplyr",
             "ggplot2",
             "knitr",
             "kableExtra",
             "RColorBrewer",
             "readxl",
             "here",
             "Hmisc",
             "scales",
              "ggpmisc",
             "ggpubr",
             "wesanderson",
             "RColorBrewer",
             "randomcoloR",
             "ggdist",
             "gmodels",
             "rcompanion",
             "gghalves",
             "ggbeeswarm",
             "rstatix",
             "multcompView",
             "ggstatsplot",
             "knitr")

## Load or install&load
package.check <- lapply(packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

## Source the functions
source(here("eLowQuantUser/eLowQuant-Functions-V20210407.R"))
```

```{r}
## Load the Data
FILE.PATH <- here("LOD_LOQ/pycno-fullSC-data-Hakai.csv")
DAT_RAW <- read.csv(FILE.PATH) %>% rename(SQ=TargetConcentration, Concentration=MeasuredConcentration) %>% 
  mutate(SQ = na_if(SQ, 0)) %>% 
  mutate(Concentration = na_if(Concentration, 0),
         PositiveDroplets = na_if(PositiveDroplets,0)) %>% 
         filter(., SQ < 312499)

  
```

## Instructions
This file performs the computations in the publication, 'A Statistical Model for Calibration and Computation of Detection and Quantification Limits for Low Copy Number Environmental DNA samples' by Lesperance, Allison, Bergman, Hocking, Helbing, *Environmental DNA*, 2021, 00, 1-12 https://doi.org/10.1002/edn3.220. Code obtained from eLowQuant v20220912.

This code has been modified by Zachary Gold as a part of the manuscript Molecular Star Gazing: Development and Validation of an Environmental DNA Assay for the Imperiled Sunflower Sea Star (*Pycnopodia helianthoides*) to be able to handle both ddPCR and qPCR data. Since this method only requires binary data (0,1) for the model fitting, we apply the eLowQuant method to ddPCR calculations using the number of positive droplets. We recognize that in the ddPCR field there are multiple schools of thought on how to set a LoB and LoD. We include a parameter of minimum number of acceptable droplets (e.g. 1,2,3) as a detection.

* Data set csv file requirements.  Columns:  Target, Lab, Cq, SQ
* For nondetects, set Cq to be empty or NA value
* For negative controls, set Sq to be empty or 0 or NA value
* Include negative controls in the csv file!
* DO NOT DUPLICATE Target names over different Labs!!

* This code uses observations with nonempty data and where phat <1 (num detect<num technical replicates)
  * Functionally this requires a minimum of 3 standard curve values where fewer than 100% of all replicates amplified. This may require including additional lower standards than are typically run on a qPCR/ddPCR plate
* Only the SQ's up to the first one with  phat==1 are used.
* Allows for variable numbers of SQ levels per Target.
* Assumes SQs == NA are zero, i.e. are negative controls.

* The code uses the R function optim.  A convergence code 0 indicates successful completion.  
* Ignore warnings if results are sensible.

* EXECUTE EACH CHUNK LOOKING AT THE OUTPUT.  IN PARTICULAR, LOOK AT THE GRAPHS IN
THE CHUNK CALLED 'PlotPois' TO DETERMINE IF THE MODEL IS APPROPRIATE.  IT IS NOT
APPROPRIATE IF 'lm Rsq' is small, i.e. near zero!

* If you wish to knit to pdf AND you do NOT have a version of Latex installed
on your computer, then run the following in your RStudio console:  
install.packages(“tinytex”); tinytex::install_tinytex()



```{r READIN, include=FALSE}
## READIN paragraph adapted from Merkes, Christopher <cmerkes@usgs.gov>
##   Used only to read in the data.


DAT <- DAT_RAW %>% 
  mutate(., Lab = "Hakai ddPCR") %>% 
  mutate(., Cq=PositiveDroplets)

## Check the data:
if(sum(colnames(DAT)=="Target")!=1) { #Is there a "Target" column?
  A <- grep("target", colnames(DAT), ignore.case=T)
  if(length(A)==1) { colnames(DAT)[A] <- "Target" } #Rename target column if it is mispelled but can be identified and there is only 1.
  if(length(A)!=1) { write("There is a problem with the 'Target' column.\n\n",file="Analysis Log.txt",append=T) } #Add error message to analysis log.
  if(length(A)>1) { cat("ERROR: multiple 'Target' columns detected.",colnames(DAT)[A],sep="\n") }
  if(length(A)==0) { print("ERROR: cannot detect 'Target' column.") }
}

if(sum(colnames(DAT)=="Lab")!=1) { #Is there a "Lab" column?
  A <- grep("lab", colnames(DAT), ignore.case=T)
  if(length(A)==1) { colnames(DAT)[A] <- "Lab" } #Rename Lab column if it is mispelled but can be identified and there is only 1.
  if(length(A)!=1) { write("There is a problem with the 'Lab' column.\n\n",file="Analysis Log.txt",append=T) } #Add error message to analysis log.
  if(length(A)>1) { cat("ERROR: multiple 'Lab' columns detected.",colnames(DAT)[A],sep="\n") }
  if(length(A)==0) { print("ERROR: cannot detect 'Lab' column.") }
}

if(sum(colnames(DAT)=="Cq")!=1) { #Is there a "Cq" column?
  A <- grep("cq|ct|cycle",colnames(DAT),ignore.case=T)
  if(length(A)==1) { colnames(DAT)[A] <- "Cq" } #Rename cq column if it is mispelled but can be identified and there is only 1.
  if(length(A)!=1) { write("There is a problem with the 'Cq' column.\n\n",file="Analysis Log.txt",append=T) } #Add error message to analysis log.
  if(length(A)>1) { cat("ERROR: multiple 'Cq' columns detected.",colnames(DAT)[A],sep="\n") }
  if(length(A)==0) { print("ERROR: cannot detect 'Cq' column.") }
}

if(sum(colnames(DAT)=="SQ")!=1) { #Is there a "SQ" column?
  A <- grep("sq|copies|starting|quantity",colnames(DAT),ignore.case=T)
  if(length(A)==1) { colnames(DAT)[A] <- "SQ" } #Rename SQ column if it is mispelled but can be identified and there is only 1.
  if(length(A)!=1) { write("There is a problem with the 'SQ' column.\n\n",file="Analysis Log.txt",append=T) } #Add error message to analysis log.
  if(length(A)>1) { cat("ERROR: multiple 'SQ' columns detected.",colnames(DAT)[A],sep="\n") }
  if(length(A)==0) { print("ERROR: cannot detect 'SQ' column.") }
}

## Ensure data is in the proper format:
DAT$Target <- as.factor(DAT$Target)
DAT$Lab <- as.factor(DAT$Lab)  #ML
DAT$Cq <- suppressWarnings(as.numeric(as.character(DAT$Cq))) #Non-numerical values (i.e. negative wells) will be converted to NAs
DAT$SQ <- suppressWarnings(as.numeric(as.character(DAT$SQ))) #Non-numerical values (i.e. NTC) will be converted to NAs

## ML Remove positive controls
# dim(DAT)
# DAT <- DAT[DAT$Content!='Pos Ctrl',]
# dim(DAT)

## ML assume SQs == NA are zero - negative controls
DAT$SQ[is.na(DAT$SQ)] <- 0
summary(DAT) %>% kable(digits=2)
DAT.df <- data.frame(DAT)
DAT.df %>%  kable(format="pandoc", digits=3)
```


### Process/Summarize samples by Lab; Compute the Poisson estimates of SQ

Hindson et al "High-Throughput Droplet Digital PCR System for Absolute 
Quantitation of DNA Copy Number", Anal. Chem., 2011, 83 (22), pp 8604–8610
use a Poisson approximation for quantitation.
Before that, Dube et al. 2008, "Mathematical analysis of copy number 
variation in a DNA sample
using digital PCR on a nanofluidic device", PloS One, Vol 3, Issue 8, e2876,
model the number of molecules in each chamber as a Poisson process, giving the 
relationship between $p$ and $\lambda$.

```{r detect, warnings=FALSE}
## Summarize data by (Target, SQ), detect=#detections, n=#tech reps
DAT.Tar.SQ <-  DAT.df %>%
  group_by(Target, SQ) %>%
  summarise(detect=sum(!is.na(Cq)), n=n(),  Cqmean=mean(Cq, na.rm=TRUE), 
            Lab=Lab[1] )
DAT.Tar.SQ <- droplevels(data.frame(DAT.Tar.SQ))
#dim(DAT.Tar.SQ); cat('dim, before negative controls added')
#summary(DAT.Tar.SQ) %>% kable(digits=2)

uLabs <- unique(DAT.Tar.SQ$Lab) #unique labs

DAT.Tar.SQ <- arrange(DAT.Tar.SQ, Lab, Target, SQ) #sort data by SQ in Target in Lab

## Add variables to data set:  L10.SQ, phat, ... 
DAT.Tar.SQ <- within(DAT.Tar.SQ, {
  L10.SQ <- log10(SQ)  
  phat <- detect/n           #sample proportion detect
  vphat <- phat*(1-phat)/n   #var of phat
  lamhat <- -log(1-phat) 
  vlamhat <- phat/n/(1-phat)  #var of lamhat using the delta method
  sdlamhat <- sqrt(vlamhat)   #sd of lamhat using the delta method
  MElamhat <- 1.96*sdlamhat  #margin of error for lambda hat using delta method
}
)



## All Targets and Labs **DO NOT DUPLICATE Target names over Labs!!
uLabs <- unique(DAT.Tar.SQ$Lab)
uTargets <- unique(DAT.Tar.SQ$Target)
nTargets <- length(uTargets)
uLabsTargets <- unique(DAT.Tar.SQ[,c('Lab','Target')])
uLabsTargets$Lab <- as.character(uLabsTargets$Lab)
#ensure ulabsTargets in same order as uTargets
uLabsTargets <- uLabsTargets[match(uLabsTargets$Target, uTargets),]  
uLabsTargets.names <- apply(uLabsTargets, 1, paste, collapse=', ')

## Print summary tables.  If sink.indicator set to TRUE, writes tables to file.

knitr::kable(DAT.Tar.SQ, digits=3, caption=paste(uLabsTargets$Lab,uLabsTargets$Target), format = "markdown")


```



```{r ExactTransCI}
## Exact 95% Binomial Confidence intervals => backtransform given alpha and beta
## Binomial bound from Julious 2005, Stat in Medicine

DAT.Tar.SQ <- within(DAT.Tar.SQ, {
  CIexphat.lower <-  1 - qbeta(.975, n-detect+1, detect)  #exact phat bounds
  CIexphat.upper <-  qbeta(.975, detect+1, n-detect)

## Use transformed exact phat bounds
  Lamhatex.Lower <- -log(1 - CIexphat.lower)
  Lamhatex.Upper <- -log(1 - CIexphat.upper)
}
)

```

### Plot the Poisson estimates (and CI) of SQ for levels that had non-detects
Only the first levels of SQ that had non-detects are analyzed. 
Red line is least squares linear regression line.

LOOK AT THE GRAPHS IN
THE CHUNK CALLED 'PlotPois' TO DETERMINE IF THE MODEL IS APPROPRIATE.  IT IS NOT
APPROPRIATE IF 'lm Rsq' is small, i.e. near zero!

*ML* will error if all phats==1

```{r PlotPois, warning=FALSE}


## Use observations with nonempty data and where phat <1
## Only the SQ's before the first one with phat==1 are used
## Allows for variable numbers of SQ levels per Target

nndetect <- vector("list", nTargets) 
nrowTarget <- rep(0, length=nTargets)

for(i in 1:nTargets) {
  Target.dat <- subset(DAT.Tar.SQ, Target==uTargets[i])
  bSQ <- !is.na(Target.dat$phat)  
  lastSQ <- as.logical(cumprod(Target.dat$phat!=1 & bSQ)) 
## removes first observations with SQ with phat=1 and larger SQs
  Target.dat <- Target.dat[lastSQ,]
  nndetect[i] <- list(Target.dat )
  nrowTarget[i] <- nrow(Target.dat)
 
  if(nrow(nndetect[[i]]) < 2) {cat(paste('Too few values for ', uTargets[i])); next}
  
  maxSQ <- max(Target.dat$SQ)
  maxlamhat <- max(Target.dat$lamhat)
  plot(Target.dat$SQ, Target.dat$lamhat, xlog=TRUE, ylab='Lambda hat',
       xlab='Starting copy number',
       ylim=c(0, maxlamhat), xlim=c(0, maxSQ), main=uLabsTargets.names[i])
## Transformed Exact CI
  arrows(Target.dat$SQ, Target.dat$Lamhatex.Lower, Target.dat$SQ, 
         Target.dat$Lamhatex.Upper,
         length=0.05, angle=90, code=3)
## overlay simple regression line and R-squared
  jlm <- lm(lamhat ~ SQ, data=Target.dat)
  abline(jlm, col=2)
  legend("topleft", paste('lm Rsq=',round(summary(jlm)$r.squared, 2)), bty="n")
  cat("\n\n")
}


```

\pagebreak

Both the intercept and no intercept models are fit to the data.  The 'best' of
the two models is the one with the largest Likelihood Ratio test p-value.
The 'best' model will be identified in the chunk called *Manuscript*.



## Intercept Model

### Estimate Poisson models - intercept model 

```{r MLEfits, warning=FALSE, fig.height=4.5, fig.width=6.5}

# List of results
Calib.fit.estimates <- vector("list", nTargets)  #list of fit estimates
names(Calib.fit.estimates) <- uTargets
Calib.fit.all <- vector("list", nTargets)  #list of fits all
Calib.fit.LLRp <- vector("numeric", nTargets)
names(Calib.fit.LLRp) <- uTargets
Calib.fit.res <- matrix(0, nrow=nTargets, ncol=4)
rownames(Calib.fit.res) <- uTargets
colnames(Calib.fit.res) <- c("convergence", "LLR", "degf","Pval")
Calib.fit.res <- data.frame(Calib.fit.res)

for(i in 1:nTargets){
  if(nrow(nndetect[[i]]) < 3) {Calib.fit.estimates[[i]] <- NULL
     cat(paste('Too few values for ', uTargets[i]), '\n')
    next}
  Target.dat <- nndetect[[i]]
  jlm <- lm(lamhat ~ SQ, data=Target.dat) #starting values for alpha and beta

  Calib.fit <- optim(par=pmax(c(0.01, 0.01), coef(jlm)), fn=Calib.LLik,  
                     nd=Target.dat$detect,
                     S=Target.dat$SQ, nn=Target.dat$n, gr=Calib.dLLik,
                     method="BFGS", control=list(fnscale=-1),  hessian=TRUE)

  Calib.fit.all[[i]] <- Calib.fit
  if(nrow(Target.dat)==2) {
    Calib.fit.Var <- matrix(0, 2, 2)  #singular hessian for 2 observations
    }  else { 
    Calib.fit.Var <- solve(-Calib.fit$hessian)
    }
  cmat <- cbind(Calib.fit$par, sqrt(diag(Calib.fit.Var)))
  cmat <- cbind(cmat, cmat[,1]/cmat[,2])
  cmat <- cbind(cmat, 2*pnorm(-cmat[, 3]))
  colnames(cmat) <- c("Estimate", "Std.Err", "Z value", "Pr(>z)")
  if (nrow(cmat)==1) {rownames(cmat) <- c("beta")
  } else  rownames(cmat) <- c("alpha","beta")
  Calib.fit.estimates[[i]] <- cmat
  
  #Likelihood ratio statistic and pvalue for goodnes-of-fit of the model
  Calib.degf <- nrow(Target.dat) - length(Calib.fit$par)
  Calib.LLR <- 2*(Bin.LLik(Target.dat$detect, Target.dat$n) 
                  - Calib.fit$value)  
  Calib.LLR.pv <- pchisq(Calib.LLR, Calib.degf, lower.tail = FALSE)
  Calib.fit.LLRp[i] <- Calib.LLR.pv
  Calib.fit.res[i,] <- c(Calib.fit$convergence, Calib.LLR, Calib.degf, Calib.LLR.pv)
  
  #Compute fitted values for ML model
  Calib.fitted <- Calib.fit$par[1] + Calib.fit$par[2]*  Target.dat$SQ  

  # Plot calibration curve on lambda scale
  maxSQ <- max(Target.dat$SQ)
  maxlamhat <- max(Target.dat$lamhat)

  plot(Target.dat$SQ, Target.dat$lamhat, xlog=TRUE, 
       ylab='Lambda hat', xlab='Starting copy number',
       ylim=c(0, maxlamhat), xlim=c(0, maxSQ),  las=1,
       main=paste0("Intercept Model: Calibration Curve on Lambda Scale ", uTargets[i]))
  abline(Calib.fit$par[1], Calib.fit$par[2], col=4)
  legend("topleft", legend=c('ML fit'), lty=1, col=4, bty="n")
  arrows(Target.dat$SQ, Target.dat$Lamhatex.Lower, Target.dat$SQ, 
         Target.dat$Lamhatex.Upper,
         length=0.05, angle=90, code=3)
 
  # Plot calibration curve on phat scale
  # Compute minSQ such that phat~=1
  maxSQa <-max( -(Calib.fit$par[1] + log(.01))/Calib.fit$par[2], maxSQ)
  sqs <- seq(0, maxSQa, by=.1)
  Calib.phat <- 1 - exp(-(Calib.fit$par[1] + Calib.fit$par[2] * sqs))

  plot(sqs, Calib.phat, xlog=TRUE, 
       ylab='Proportion detect', xlab='Starting copy number', type='l', col=4, las=1, 
       ylim=c(0, 1), xlim=c(0, maxSQa),  main=paste0("Intercept Model: Calibration Curve on Phat Scale ", uTargets[i]))
  points(Target.dat$SQ, Target.dat$phat)
  legend("topleft", legend=c('ML fit'), lty=1, col=4, bty="n")

  cat(as.character(uTargets[i]), "\n")
  cat("Convergence=", Calib.fit$convergence, "\n")
  printCoefmat(cmat, digits=3)
  cat('LLR test stat=', Calib.LLR, ', df= ', Calib.degf, ', p-value=', Calib.LLR.pv, "\n\n")

}   

  
``` 
**Convergence of 0 indicates a successful model fit**

\newpage
